---
title: "Assignment 4"
author: "Charuvi Begwani, Aijing Li"
date: "2/17/2022"
output: html_document
---

#Networks - Existing Scenario
This file determines the existing characteristics of the street and transit networks within Salt Lake MSA.

#Load relevant libraries
```{r}
library(tidyverse)
library(sf)
library(tigris)
library(osmdata)
library(devtools)
library(tidytransit)
library(here)
```


#Download a GTFS feed
We downloaded data on Utah Transit Authority in the GTFS format from OpenMobilityData which maintains an archive of GTFS feeds for cities throughout the world. We downloaded the GTFS feed dated 15 February 2022 for our study area from here (https://transitfeeds.com/p/utah-transportation-authority) and saved the zipped file to the networks subfolder in the existing folder of our project directory. 

#Download an OpenStreetMap network
We represent the existing roadway network in a format that to allow us to calculate travel times. OpenStreetMaps is helpful for this.
```{r}
# Load the MSA boundaries
boundary <- core_based_statistical_areas() %>%
  filter(GEOID == "41620")

# Define a bounding box containing the MSA
saltlake_bbox <- st_bbox(boundary)

q <- opq(bbox = saltlake_bbox) %>% # create a query
  add_osm_feature(key = 'highway') %>% # request only road data
  osmdata_xml(file = 'existing/networks/streets.osm') # download osm file
```


#Convert an OSM file to a PBF file
You will be downloading OpenStreetMap data (and potentially editing it) in the .osm file format. To use it with the the the r5r package, you’ll need to convert it to a .pbf file (which is just a compressed version of the *.osm file).

#Create zone centroids
To be able to calculate the travel times to and from each centroid, we generate a set of centroid points from the zone boundary.
```{r}
centroids <- here("zones",
                  "boundaries.geojson") %>%
  st_read() %>%
  st_centroid() %>%
  st_transform("WGS84") %>%
  rename(id = GEOID)
```

#Install scenRios package
Install the scenRios package from Github developed by Dr. Carole Voulgaris to edit the the transit network or the street network.
```{r}
install_github("https://github.com/c-voulgaris/scenRios")
```


#Generate skims
A skim is a matrix of travel times (or distances or costs) for all possible origin-destination pairs in your model. The R5R package is really useful for generating skims.


