---
title: "Assignment 3"
author: "Aijing Li, Charuvi Begwani"
date: "2/4/2022"
output: html_document
---


```{r,warning=FALSE,message=FALSE}
library(tibble)
library(tidycensus)
library(sf)
library(RColorBrewer)
library(ggthemes)
library(dplyr)
library(treemapify)
library(tidyverse)
library(knitr)
library(units)

```



# Introduction 
Main Street is one of the key commercial streets in Salt Lake City with banks, major retailers and high pedestrian footfall. In a few weeks in September and October 2020, a block on the Main Street was shut down for cars on the weekends (Thursday - Saturday) from 6-10 p.m. Referred to as the Downtown SLC Open Streets program, this was done in an attempt to revitalize the downtown area and encourage patrons. The new spaces created, and outdoor dining and retail services initiated as a result of social distancing requirements due to the pandemic, were leveraged to “expand options for pedestrians” and benefit on-street businesses. In response to the success of this initiative, it was repeated the next year in April with the concerned stretch expanded to include South Temple to 400 South. As before, this was an attempt to invite pedestrians with some ongoing activities and support local businesses. This news article reports that Downtown Alliance, the group leading this, is considering asking city to make it a permanent, year-round “pedestrian zone.” 

We are interested in modelling Main Street pedestrianization and closure for cars to explore how impactful the plan for a permanent pedestrian zone may be. We think this is especially relevant since Open Streets have been implemented all across the country, post-pandemic, and many other cities are also considering similar pedestrianization projects with the success of outdoor dining and retail activities.

#Getting census variable names
We first retrieve census data on population characteristics through the American Commmunity Survey 2019, using the tidycensus package.

```{r}
area_vars_2019 <- load_variables(2019, "acs5")
```
#Download census data.
We download census data for the Salt Lake Metropolitan Statistical Area for selected variables that are relevant to this analysis, including number of households, household size, vehicle ownership and household income. This data is available at the census block level, and we retrieve it for Salt Lake and Tooele Counties that constitute the Salt Lake MSA.

```{r}

vars <- c(total_hhs = 'B08203_001',
          no_veh = 'B08203_002',
          avg_hh_size = "B25010_001",
          
          hh_1person = 'B08201_007',
          hh_2person = 'B08201_013',
          hh_3person = 'B08201_019',
          hh_4person_plus = 'B08201_025',
          
          inc_lt_10k = 'B19001_002',
          inc_btw_10k_15k = 'B19001_003',
          inc_btw_15k_20k = 'B19001_004',
          inc_btw_20k_25k = 'B19001_005',
          inc_btw_25k_30k = 'B19001_006',
          inc_btw_30k_35k = 'B19001_007',
          inc_btw_35k_40k = 'B19001_008',
          inc_btw_40k_45k = 'B19001_009',
          inc_btw_45k_50k = 'B19001_010',
          inc_btw_50k_60k = 'B19001_011',
          inc_btw_60k_75k = 'B19001_012',
          inc_btw_75k_100k = 'B19001_013',
          inc_btw_100k_125k = 'B19001_014',
          inc_btw_125k_150k = 'B19001_015',
          inc_btw_150k_200k = 'B19001_016',
          inc_gt_200k = 'B19001_017'
          )

census <- get_acs(geography = 'tract',
                 state = 'UT',
                 county = c('Salt Lake', 'Tooele'),
                 variables = vars,
                 output = 'wide',
                 geometry = TRUE)

```
#Combine income variables to create three new income categories.
We had retreived census data on the total number of households earning incomes in various income categories at the census block level. We then collapsed and determined the tract level data for that and combined all the income categories to form five broad categories based on the regional quintiles at 20%, 35%, 60% and 80% levels. These are under 35,000; 35,000 - 60,000; 60,000 - 100,000; 100,000 - 150,000; and 150,000 - $200,000.

```{r}
census <- census %>% 
  mutate(inc_quint_20 = inc_lt_10kE+inc_btw_10k_15kE+
          inc_btw_15k_20kE+inc_btw_20k_25kE+inc_btw_25k_30kE+inc_btw_30k_35kE)%>%
  mutate(inc_quint_35 = inc_btw_35k_40kE + inc_btw_40k_45kE 
        + inc_btw_45k_50kE + inc_btw_50k_60kE)%>%
  mutate(inc_quint_60 = inc_btw_60k_75kE + inc_btw_75k_100kE)%>%
  mutate(inc_quint_80 = inc_btw_100k_125kE+inc_btw_125k_150kE)%>%
  mutate(inc_quint_100 = inc_btw_150k_200kE+inc_gt_200kE)%>%
select(GEOID, 
         total_hhsE,
         no_vehE,
         avg_hh_sizeE,
         hh_1personE,
         hh_2personE,
         hh_3personE,
         hh_4person_plusE,
        
         inc_quint_20,
         inc_quint_35,
         inc_quint_60,
         inc_quint_80,
         inc_quint_100,
        
          inc_lt_10kE,
          inc_btw_10k_15kE,
          inc_btw_15k_20kE,
          inc_btw_20k_25kE,
          inc_btw_25k_30kE,
          inc_btw_30k_35kE,
          inc_btw_35k_40kE,
          inc_btw_40k_45kE,
          inc_btw_45k_50kE,
          inc_btw_50k_60kE,
          inc_btw_60k_75kE,
          inc_btw_75k_100kE,
          inc_btw_100k_125kE,
          inc_btw_125k_150kE,
          inc_btw_150k_200kE,
          inc_gt_200kE)
```
#Download block level employment data for Utah from the LEHD dataset.
For this exercise, we first retreived employment data from the Longitudinal Employer-Household Dynamics (LEHD) dataset. We pulled census block-level data for the state of Utah from the Workplace Characteristics section for 2019. Variables we pulled include:
- Total employment or number of workers
- Basic employment which we added (includes Agriculture, Forestry,
Fishing, and Hunting (CNS01), Mining and extraction (CNS02), Utilities
(CNS03), Construction (CNS04), Manufacturing (CNS05), Wholesale
trade (CNS06) and Transportation and warehousing (CNS06).
- Retail employment
- Service employment (total - basic - retail)

```{r}

library(tidyverse)

lehd_blocks <- read_csv('https://lehd.ces.census.gov/data/lodes/LODES7/ut/wac/ut_wac_S000_JT00_2019.csv.gz') %>%
  rename(total_emp = C000) %>%
  mutate(basic_emp = CNS01+CNS02+CNS03+CNS04+CNS05+CNS06+CNS08+CNS09) %>%
  rename(retail_emp = CNS07) %>%
  mutate(service_emp = total_emp - basic_emp - retail_emp) %>%
  select(w_geocode, total_emp, basic_emp, retail_emp, service_emp)
```

#Collapse block level employment data to tract level data for Utah.
We then collapse the block level data to determined employment characteristics at the tract level for the state data. 
```{r}
library(tidyverse)

lehd_tracts <- lehd_blocks %>%
  mutate(w_geocode = as.character(w_geocode)) %>%
  mutate(GEOID = substr(w_geocode, 1, 11)) %>%
  select(-w_geocode) %>%
  group_by(GEOID) %>%
  summarize(across(everything(), ~sum(.)))
```

#Join tract level employment data with population data 
We then combine employment data for Utah state with the previously compiled population data tracts within Salt Lake MSA boundary. This filters down the state level employment data to data for our two selected counties (Salt Lake and Tooele) in the Salt Lake MSA area.

```{r}
zones <- left_join(census, lehd_tracts)

```

#Separate zone geometry or boundaries from the zone dataset
```{r}
zone_boundaries <- zones %>%
  select(GEOID, geometry)

zone_data <- zones %>%
  st_drop_geometry()
```

#Summarize tract statistics
We summarize all key statistics with regard to the tract level data and determine mean, median and standard deviation for all selected variables. We all find the total number of households and employees in our MSA.
```{r}
#
table <- zone_data %>%
  pivot_longer(cols=c(total_hhsE,no_vehE,
         inc_quint_20,
         inc_quint_35,
         inc_quint_60,
         inc_quint_80,
         inc_quint_100,
          total_emp, basic_emp, retail_emp, service_emp,
         hh_1personE, hh_2personE, hh_3personE, hh_4person_plusE),
               names_to = "Variable",
               values_to = "value") %>%
  group_by(Variable) %>%
  summarize(Average = mean(value, na.rm=TRUE),
            `Standard deviation` = sd(value, na.rm=TRUE),
            Median = median(value, median(value, na.rm=TRUE)))

write_csv(zone_data, "images/zone_summary.csv", append = FALSE)

sum(census$total_hhsE)
sum(zone_data$retail_emp)
sum(zone_data$service_emp)
sum(zone_data$total_emp)
```

# Regional Data
We find the distribution of household incomes for the region as a whole so that we can determine the quintiles. We use these same quintiles to create income categories for our tract level data.
```{r}
vars <- c(inc_lt_10k = 'B19001_002',
          inc_btw_10k_15k = 'B19001_003',
          inc_btw_15k_20k = 'B19001_004',
          inc_btw_20k_25k = 'B19001_005',
          inc_btw_25k_30k = 'B19001_006',
          inc_btw_30k_35k = 'B19001_007',
          inc_btw_35k_40k = 'B19001_008',
          inc_btw_40k_45k = 'B19001_008',
          inc_btw_45k_50k = 'B19001_010',
          inc_btw_50k_60k = 'B19001_011',
          inc_btw_60k_75k = 'B19001_012',
          inc_btw_75k_100k = 'B19001_013',
          inc_btw_100k_125k = 'B19001_014',
          inc_btw_125k_150k = 'B19001_015',
          inc_btw_150k_200k = 'B19001_016',
          inc_gt_200k = 'B19001_017')


region_inc <- get_acs(
  geography = "metropolitan statistical area/micropolitan statistical area",
  variables = vars,
  summary_var = 'B19001_001',
  geometry = FALSE) %>%
  filter(GEOID == "41620") %>%
  mutate(pct = estimate / summary_est) %>%
  mutate(cumul_pct = cumsum(pct)) %>%
  select(variable, cumul_pct)
```
```{r}
kable(region_inc, digits=2)
```



# Calculate Area & Density Variables
```{r}
census <- census %>%
  mutate(tract_area = st_area(geometry)) %>%
  mutate(tract_area = set_units(tract_area, "mi2")) 
```

```{r}
zones <- zones %>% 
  mutate(total_hh_density = total_hhsE/census$tract_area)%>% 
  mutate(no_veh_density = no_vehE/census$tract_area)%>% 
  mutate(total_emp_density = total_emp/census$tract_area)%>% 
  mutate(basic_emp_density = basic_emp/census$tract_area)%>% 
  mutate(service_emp_density = service_emp/census$tract_area)%>% 
  mutate(retail_emp_density = retail_emp/census$tract_area)
```


#Creating chloropleth maps for household, income and employment data
```{r}
total_hh_map <- ggplot(zones) +
  geom_sf(aes(fill = as.numeric(total_hh_density)), color="NA")

map_palette <- brewer.pal(5, "PuBuGn")

total_hh_map + 
theme_void(base_size = 11) +
scale_fill_gradientn(colors = map_palette,  name = "Density of households/mile^2") 

ggsave("images/total_hh_map.png",
       width = 417.67, height = 200, units = "mm",dpi = 300)
```

```{r}
no_veh_map <- ggplot(zones) +
  geom_sf(aes(fill = as.numeric(no_veh_density)), color="NA")

map_palette <- brewer.pal(5, "PuBuGn")

no_veh_map + 
theme_void(base_size = 11) +
scale_fill_gradientn(colors = map_palette,  name = "Density of zero-vehicle\nhouseholds/mile^2") 

ggsave("images/no_veh_map.png",
       width = 417.67, height = 200, units = "mm",dpi = 300)
```

```{r}
total_emp_map <- ggplot(zones) +
  geom_sf(aes(fill = as.numeric(total_emp_density)), color="NA")

map_palette <- brewer.pal(5, "PuBuGn")

total_emp_map + 
theme_void(base_size = 11) +
scale_fill_gradientn(colors = map_palette,  name = "Density of employees\ndensity/mile^2") 

ggsave("images/total_emp_map .png",
      width = 417.67, height = 200, units = "mm",dpi = 300)
```

```{r}
basic_emp_map <- ggplot(zones) +
  geom_sf(aes(fill = as.numeric(basic_emp_density)), color="NA")

map_palette <- brewer.pal(5, "PuBuGn")

basic_emp_map + 
theme_void(base_size = 11) +
scale_fill_gradientn(colors = map_palette,  name = "Density of basic\nemployees density/mile^2") 

ggsave("images/basic_emp_map.png",
      width = 417.67, height = 200, units = "mm",dpi = 300)
```


```{r}
retail_emp_map <- ggplot(zones) +
  geom_sf(aes(fill = as.numeric(retail_emp)), color="NA")

map_palette <- brewer.pal(5, "PuBuGn")

retail_emp_map + 
theme_void(base_size = 11) +
scale_fill_gradientn(colors = map_palette,  name = "Density of retail\n employees/miles^2") 

ggsave("images/retail_emp_map.png",
       width = 417.67, height = 200, units = "mm",dpi = 300)
```
```{r}
service_emp_map <- ggplot(zones) +
  geom_sf(aes(fill = service_emp), color="NA")

map_palette <- brewer.pal(5, "PuBuGn")

service_emp_map + 
theme_void(base_size = 8) +
scale_fill_gradientn(colors = map_palette,  name = "Total number of service employees") 

ggsave("images/service_emp_map.png",
      width = 417.67, height = 200, units = "mm",dpi = 300)
```
# Generate treemap diagrams for income and household size composition of study area
```{r}
income_buckets <- c(sum(census$inc_quint_20)/sum(census$total_hhsE),
                    sum(census$inc_quint_35)/sum(census$total_hhsE),
                    sum(census$inc_quint_60)/sum(census$total_hhsE),
                    sum(census$inc_quint_80)/sum(census$total_hhsE),
                    sum(census$inc_quint_100)/sum(census$total_hhsE)                    
                    ) 
income_tibble <- tibble(bucket = c("Quantile 20%: Below 35k",
                               "Quantile 35%: 35-60k",
                               "Quantile 60%: 60-100k",
                               "Quantile 80%: 100-150k",
                               "Quantile 100%: Above 150k"),

                     Percentage = as.numeric(income_buckets*100)
                     )

ggplot(income_tibble, aes(area = Percentage, fill= bucket)) +
  geom_treemap(show.legend = FALSE) +
  geom_treemap_text(aes(label = paste(bucket, "\n",
                                      prettyNum(Percentage, digits = 0),
                                      "%",sep = "")), 
                        color = "black") +
  scale_fill_brewer(palette = "PuBuGn")

ggsave("images/treemap_income.png",
      width = 20, height = 9, units = "cm",dpi = 300)
```


```{r}
hhsize_buckets <- c(sum(census$hh_1personE)/sum(census$total_hhsE),sum(census$hh_2personE)/sum(census$total_hhsE),sum(census$hh_3personE)/sum(census$total_hhsE),sum(census$hh_4person_plusE)/sum(census$total_hhsE)) 
hhsize_tibble <- tibble(bucket = c("Households with 1 person",
                               "Households with 2 person",
                               "Households with 3 person",
                               "Households with 4 person and plus"),

                     Percentage = as.numeric(hhsize_buckets*100)
                     )

ggplot(hhsize_tibble, aes(area = Percentage, fill= bucket)) +
  geom_treemap(show.legend = FALSE) +
  geom_treemap_text(aes(label = paste(bucket, "\n",
                                      prettyNum(Percentage, digits = 0),
                                      "%",sep = "")), 
                        color = "black") +
  scale_fill_brewer(palette = "PuBuGn")

ggsave("images/treemap_hhsize.png",
       width = 20, height = 9, units = "cm",dpi = 300)
```

# Create Dot-Density Map
```{r}
library(sf)

hh_1person_pts <- st_sample(census, 
                            size = ceiling(census$hh_1personE/100))

hh_2person_pts <- st_sample(census, 
                            size = ceiling(census$hh_2personE/100))

hh_3person_pts <- st_sample(census, 
                            size = ceiling(census$hh_3personE/100))

hh_4person_pts <- st_sample(census, 
                            size = ceiling(census$hh_4person_plusE/100))
```

```{r}
hh_1person_df <- tibble(hh_size = rep("1 person", 
                                      length(hh_1person_pts))) %>%
  st_sf(geom = hh_1person_pts)
                        
hh_2person_df <- tibble(hh_size = rep("2 people", 
                                      length(hh_2person_pts))) %>%
  st_sf(geom = hh_2person_pts)
 
hh_3person_df <- tibble(hh_size = rep("3 people", 
                                      length(hh_3person_pts))) %>%
  st_sf(geom = hh_3person_pts)

hh_4person_df <- tibble(hh_size = rep("4 or more people", 
                                      length(hh_4person_pts))) %>%
  st_sf(geom = hh_4person_pts)

hh_size_pts <- rbind(hh_1person_df, hh_2person_df, hh_3person_df, hh_4person_df)
```

```{r}
ggplot(census) + 
  geom_sf(color = "white") +
  geom_sf(data = hh_size_pts, 
          aes(color = hh_size), 
          alpha = 0.3,
          size = 0.1) +
  scale_color_brewer("Household size\n(each points represents\n100 households)",
                     palette = "Set1") +
  theme_void()  +
  guides(color = guide_legend(override.aes = list(size=8, alpha = 0.6)))

ggsave("images/dotdensitymap_hhsize.png",
       width = 40, height = 20, units = "cm",dpi = 300)
```
```{r}
inc_quint_20_pts <- st_sample(census, 
                            size = ceiling(census$inc_quint_20/100))

inc_quint_35_pts <- st_sample(census, 
                            size = ceiling(census$inc_quint_35/100))

inc_quint_60_pts <- st_sample(census, 
                            size = ceiling(census$inc_quint_60/100))

inc_quint_80_pts <- st_sample(census, 
                            size = ceiling(census$inc_quint_80/100))

inc_quint_100_pts <- st_sample(census, 
                            size = ceiling(census$inc_quint_100/100))
```


```{r}
inc_quint_20_df <- tibble(inc_category = rep("Under 35,000 ", 
                                      length(inc_quint_20_pts))) %>%
  st_sf(geom = inc_quint_20_pts)
                        
inc_quint_35_df <- tibble(inc_category = rep("35,000 - 60,000", 
                                      length(inc_quint_35_pts))) %>%
  st_sf(geom = inc_quint_35_pts)
 
inc_quint_60_df <- tibble(inc_category = rep("60,000 - 100,000", 
                                      length(inc_quint_60_pts))) %>%
  st_sf(geom = inc_quint_60_pts)

inc_quint_80_df <- tibble(inc_category = rep("100,000 - 150,000", 
                                      length(inc_quint_80_pts))) %>%
  st_sf(geom = inc_quint_80_pts)

inc_quint_100_df <- tibble(inc_category = rep("150,000 - $200,000", 
                                      length(inc_quint_100_pts))) %>%
  st_sf(geom = inc_quint_100_pts)

inc_quint_pts <- rbind(inc_quint_20_df, inc_quint_35_df, inc_quint_60_df, inc_quint_80_df, inc_quint_100_df)
```

```{r}
ggplot(census) + 
  geom_sf(color = "white") +
  geom_sf(data = inc_quint_pts, 
          aes(color = inc_category), 
          alpha = 0.3,
          size = 0.1) +
  scale_color_brewer("Household income\n(each points represents\n100 households)",
                     palette = "Set1") +
  theme_void()  +
  guides(color = guide_legend(override.aes = list(size=8, alpha = 0.6)))

ggsave("images/dotdensitymap_income.png",
       width = 40, height = 20, units = "cm",dpi = 300)
```


#Save and write datasets to a file.
```{r}
st_write(zone_boundaries, "zones/boundaries.geojson", append = FALSE)
write_csv(zone_data, "existing/data/zone_data.csv", append = FALSE)
write_csv(zone_data, "alternative/data/zone_data.csv", append = FALSE)
write_csv(region_inc, "existing/data/regional_income_quintile.csv", append = FALSE)
```
```{r}
library(leaflet)
library(htmlwidgets)

map <- leaflet(census) %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(popup = ~GEOID,
              weight = 1,
              opacity = 1,
              highlightOptions =
                highlightOptions(fillColor = "red"))

saveWidget(map, file = 'zones/boundaries.html')
```



