---
title: "Assignment 10"
author: "Charuvi Begwani, Aijing Li"
date: "4/11/2022"
output: html_document
---

#Load libraries

```{r}
options(java.parameters = "-Xmx2G")

library(here)
library(tidyverse)
library(stplanr)
library(r5r)
library(sf)
library(tidyverse)
library(RColorBrewer)
```

#Load data from previous assignments including production and attraction zone IDs, the transit routes included in that transit trip, and the number of trips by each mode for each trip purpose for both existing and alternative.

```{r}
skims <- here("existing",
                     "data",
                     "skims.csv") %>%
  read_csv(
    #show_col_types = FALSE
    ) %>%
  select(fromId, 
         toId, 
         n_transit_HBO,
         n_SOV_HBO,
         n_HOV_HBO,
         n_walk_HBO,
         n_bike_HBO,
         n_transit_HBW,
         n_SOV_HBW,
         n_HOV_HBW,
         n_walk_HBW,
         n_bike_HBW,
         n_transit_NHB,
         n_SOV_NHB,
         n_HOV_NHB,
         n_walk_NHB,
         n_bike_NHB,
         routes) %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId))

skims_alt <- here("alternative",
                     "data",
                     "skims_alt.csv") %>%
  read_csv(
    #show_col_types = FALSE
    ) %>%
  select(fromId, 
         toId, 
         n_transit_HBO,
         n_SOV_HBO,
         n_HOV_HBO,
         n_walk_HBO,
         n_bike_HBO,
         n_transit_HBW,
         n_SOV_HBW,
         n_HOV_HBW,
         n_walk_HBW,
         n_bike_HBW,
         n_transit_NHB,
         n_SOV_NHB,
         n_HOV_NHB,
         n_walk_NHB,
         n_bike_NHB,
         routes) %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId))
```

#Convert P-A matrix to O-D matrix: HBO Trips

##Convert data frame to matrix
```{r}
HBO_PA_mat <- skims %>%
  od_to_odmatrix(attrib = "n_transit_HBO", 
                 name_orig = "fromId",
                 name_dest = "toId") 

HBO_PA_mat <- HBO_PA_mat[,row.names(HBO_PA_mat)]

##Transpose trip matrix
HBO_PA_mat_trans <- t(HBO_PA_mat)

##Average matrix with its transpose 
HBO_OD_mat <- (HBO_PA_mat + HBO_PA_mat_trans) / 2

##Convert matrix to data frame
HBO_OD_table <- HBO_OD_mat %>%
  odmatrix_to_od() %>%
  rename(fromId = orig,
         toId = dest) %>%
  left_join(skims) %>%
  select(-n_transit_HBO)
```
#Convert P-A matrix to O-D matrix: HBW Trips

```{r}
##Convert data frame to matrix
HBW_PA_mat <- skims %>%
  od_to_odmatrix(attrib = "n_transit_HBW", 
                 name_orig = "fromId",
                 name_dest = "toId") 

HBW_PA_mat <- HBW_PA_mat[,row.names(HBW_PA_mat)]

##Transpose trip matrix
HBW_PA_mat_trans <- t(HBW_PA_mat)

##Average matrix with its transpose and convert matrix to data frame
HBW_OD_mat <- (HBW_PA_mat + HBW_PA_mat_trans) / 2

HBW_OD_table <- HBW_OD_mat %>%
  odmatrix_to_od() %>%
  rename(fromId = orig,
         toId = dest) %>%
  left_join(skims) %>%
  select(-n_transit_HBW)

```
#Convert P-A matrix to O-D matrix: NHB Trips

```{r}
##Convert data frame to matrix
NHB_PA_mat <- skims %>%
  od_to_odmatrix(attrib = "n_transit_NHB", 
                 name_orig = "fromId",
                 name_dest = "toId") 

NHB_PA_mat <- NHB_PA_mat[,row.names(NHB_PA_mat)]

##Transpose trip matrix
NHB_PA_mat_trans <- t(NHB_PA_mat)

##Average matrix with its transpose and convert matrix to data frame
NHB_OD_mat <- (NHB_PA_mat + NHB_PA_mat_trans) / 2

NHB_OD_table <- NHB_OD_mat %>%
  odmatrix_to_od() %>%
  rename(fromId = orig,
         toId = dest) %>%
  left_join(skims) %>%
  select(-n_transit_NHB)

```

#Count trips using each route
```{r}
route_trips_HBO <- HBO_OD_table %>%
  filter(flow > 0 & !is.na(routes)) %>%
  mutate(route_1 = str_split_fixed(routes, "\\|", 3)[,1],
         route_2 = str_split_fixed(routes, "\\|", 3)[,2],
         route_3 = str_split_fixed(routes, "\\|", 3)[,3]) %>%
  pivot_longer(cols = c(route_1, route_2, route_3),
               values_to = "route") %>%
  filter(route != "") %>%
  select(route, flow) %>%
  group_by(route) %>%
  summarize(ridership_HBO = round(sum(flow)))

route_trips_HBW <- HBW_OD_table %>%
  filter(flow > 0 & !is.na(routes)) %>%
  mutate(route_1 = str_split_fixed(routes, "\\|", 3)[,1],
         route_2 = str_split_fixed(routes, "\\|", 3)[,2],
         route_3 = str_split_fixed(routes, "\\|", 3)[,3]) %>%
  pivot_longer(cols = c(route_1, route_2, route_3),
               values_to = "route") %>%
  filter(route != "") %>%
  select(route, flow) %>%
  group_by(route) %>%
  summarize(ridership_HBW = round(sum(flow)))

route_trips_NHB <- NHB_OD_table %>%
  filter(flow > 0 & !is.na(routes)) %>%
  mutate(route_1 = str_split_fixed(routes, "\\|", 3)[,1],
         route_2 = str_split_fixed(routes, "\\|", 3)[,2],
         route_3 = str_split_fixed(routes, "\\|", 3)[,3]) %>%
  pivot_longer(cols = c(route_1, route_2, route_3),
               values_to = "route") %>%
  filter(route != "") %>%
  select(route, flow) %>%
  group_by(route) %>%
  summarize(ridership_NHB = round(sum(flow)))

```

## join all purposes - existing

```{r}
#join all table and sum 
route_trip_by_purpose<-left_join(route_trips_HBO,route_trips_HBW,by="route")%>%
                        left_join(route_trips_NHB,by="route")%>%
                        select(route,ridership_HBO,ridership_HBW,ridership_NHB)%>%
                        mutate(sum_ridership=ridership_HBO+ridership_HBW+ridership_NHB)

```

## Transit ridership estimation for alternative

#Convert P-A matrix to O-D matrix: HBO Trips - alternative
##Convert data frame to matrix
```{r}
HBO_PA_mat_alt <- skims_alt %>%
  od_to_odmatrix(attrib = "n_transit_HBO", 
                 name_orig = "fromId",
                 name_dest = "toId") 

HBO_PA_mat_alt <- HBO_PA_mat_alt[,row.names(HBO_PA_mat_alt)]

##Transpose trip matrix
HBO_PA_mat_trans_alt <- t(HBO_PA_mat_alt)

##Average matrix with its transpose 
HBO_OD_mat_alt <- (HBO_PA_mat_alt + HBO_PA_mat_trans_alt) / 2

##Convert matrix to data frame
HBO_OD_table_alt <- HBO_OD_mat_alt %>%
  odmatrix_to_od() %>%
  rename(fromId = orig,
         toId = dest) %>%
  left_join(skims_alt) %>%
  select(-n_transit_HBO)
```

#Convert P-A matrix to O-D matrix: HBW Trips - alternative

```{r}
##Convert data frame to matrix
HBW_PA_mat_alt <- skims_alt %>%
  od_to_odmatrix(attrib = "n_transit_HBW", 
                 name_orig = "fromId",
                 name_dest = "toId") 

HBW_PA_mat_alt <- HBW_PA_mat_alt[,row.names(HBW_PA_mat_alt)]

##Transpose trip matrix
HBW_PA_mat_trans_alt <- t(HBW_PA_mat_alt)

##Average matrix with its transpose and convert matrix to data frame
HBW_OD_mat_alt <- (HBW_PA_mat_alt + HBW_PA_mat_trans_alt) / 2

HBW_OD_table_alt <- HBW_OD_mat_alt %>%
  odmatrix_to_od() %>%
  rename(fromId = orig,
         toId = dest) %>%
  left_join(skims_alt) %>%
  select(-n_transit_HBW)

```

#Convert P-A matrix to O-D matrix: NHB Trips - alternative

```{r}
##Convert data frame to matrix
NHB_PA_mat_alt <- skims_alt %>%
  od_to_odmatrix(attrib = "n_transit_NHB", 
                 name_orig = "fromId",
                 name_dest = "toId") 

NHB_PA_mat_alt <- NHB_PA_mat_alt[,row.names(NHB_PA_mat_alt)]

##Transpose trip matrix
NHB_PA_mat_trans_alt <- t(NHB_PA_mat_alt)

##Average matrix with its transpose and convert matrix to data frame
NHB_OD_mat_alt <- (NHB_PA_mat_alt + NHB_PA_mat_trans_alt) / 2

NHB_OD_table_alt <- NHB_OD_mat_alt %>%
  odmatrix_to_od() %>%
  rename(fromId = orig,
         toId = dest) %>%
  left_join(skims_alt) %>%
  select(-n_transit_NHB)

```

#Count trips using each route - alternative

```{r}
route_trips_HBO_alt <- HBO_OD_table_alt %>%
  filter(flow > 0 & !is.na(routes)) %>%
  mutate(route_1 = str_split_fixed(routes, "\\|", 3)[,1],
         route_2 = str_split_fixed(routes, "\\|", 3)[,2],
         route_3 = str_split_fixed(routes, "\\|", 3)[,3]) %>%
  pivot_longer(cols = c(route_1, route_2, route_3),
               values_to = "route") %>%
  filter(route != "") %>%
  select(route, flow) %>%
  group_by(route) %>%
  summarize(ridership_HBO = round(sum(flow)))

route_trips_HBW_alt <- HBW_OD_table_alt %>%
  filter(flow > 0 & !is.na(routes)) %>%
  mutate(route_1 = str_split_fixed(routes, "\\|", 3)[,1],
         route_2 = str_split_fixed(routes, "\\|", 3)[,2],
         route_3 = str_split_fixed(routes, "\\|", 3)[,3]) %>%
  pivot_longer(cols = c(route_1, route_2, route_3),
               values_to = "route") %>%
  filter(route != "") %>%
  select(route, flow) %>%
  group_by(route) %>%
  summarize(ridership_HBW = round(sum(flow)))

route_trips_NHB_alt <- NHB_OD_table_alt %>%
  filter(flow > 0 & !is.na(routes)) %>%
  mutate(route_1 = str_split_fixed(routes, "\\|", 3)[,1],
         route_2 = str_split_fixed(routes, "\\|", 3)[,2],
         route_3 = str_split_fixed(routes, "\\|", 3)[,3]) %>%
  pivot_longer(cols = c(route_1, route_2, route_3),
               values_to = "route") %>%
  filter(route != "") %>%
  select(route, flow) %>%
  group_by(route) %>%
  summarize(ridership_NHB = round(sum(flow)))

```


## join all purposes - alternative

```{r}
#join all table and sum 
route_trip_by_purpose_alt<-left_join(route_trips_HBO_alt,route_trips_HBW_alt,by="route")%>%
                        left_join(route_trips_NHB_alt,by="route")%>%
                        select(route,ridership_HBO,ridership_HBW,ridership_NHB)%>%
                        mutate(sum_ridership_alt=ridership_HBO+ridership_HBW+ridership_NHB)

```

#Find change in ridership between existing and alternative scenario
```{r}
route_ridership_by_purpose_alt <- route_trip_by_purpose_alt %>%
  rename(ridership_HBW_alt = ridership_HBW,
         ridership_HBO_alt = ridership_HBO,
         ridership_NHB_alt = ridership_NHB)

ridership_change <- left_join(route_trip_by_purpose, route_ridership_by_purpose_alt, by = "route") %>% 
  mutate(change_HBW = ridership_HBW_alt - ridership_HBW,
         change_HBO = ridership_HBO_alt - ridership_HBO,
         change_NHB = ridership_NHB_alt - ridership_NHB,
         change_total = sum_ridership_alt - sum_ridership)

ridership_change

write_csv(ridership_change, here("alternative",
              "data",
              "ridership_change.csv"))
```

#Find VMT and PMT by mode - existing
##Calculate trip distances

```{r}
centroids <- here("zones",
                   "centroids.geojson") %>%
  st_read() %>%
  filter(!st_is_empty(.))

origins <- centroids %>%
  slice(rep(1:n(), each = n())) %>%
  mutate(dest_order =
           rep(seq(1:length(centroids$id)),
               length(centroids$id)))

destinations <- origins %>%
  arrange(dest_order)
```

```{r}
# r5r_core_existing <- here("existing",
#                           "networks") %>%
#   setup_r5(verbose = FALSE)
# 
# ped_dist <- detailed_itineraries(r5r_core_existing,
#                                  origins = origins,
#                                  destinations = destinations,
#                                  mode = "WALK",
#                                  verbose = FALSE)  %>%
#   mutate(miles = distance / 1609.34) %>%
#   select(fromId, toId, miles) %>%
#   st_drop_geometry()
# 
# drive_dist <- detailed_itineraries(r5r_core_existing,
#                                  origins = origins,
#                                  destinations = destinations,
#                                  mode = "CAR",
#                                  verbose = FALSE) %>%
#   mutate(miles = distance / 1609.34) %>%
#   select(fromId, toId, miles) %>%
#   st_drop_geometry()
# 
# bike_dist <- detailed_itineraries(r5r_core_existing,
#                                  origins = origins,
#                                  destinations = destinations,
#                                  mode = "BICYCLE",
#                                  verbose = FALSE) %>%
#   mutate(miles = distance / 1609.34) %>%
#   select(fromId, toId, miles) %>%
#   st_drop_geometry()
# 
# stop_r5()
```

```{r}
bike_dist <- here("existing",
                     "data",
                     "bike_dist.csv") %>%
  read_csv(
    #show_col_types = FALSE
    ) %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId))

drive_dist <- here("existing",
                     "data",
                     "drive_dist.csv") %>%
  read_csv(
    #show_col_types = FALSE
    )  %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId))

ped_dist <- here("existing",
                     "data",
                     "ped_dist.csv") %>%
  read_csv(
    #show_col_types = FALSE
    ) %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId))
```

## Calculate PMT by mode

## Car

```{r}
PMT_car <- skims %>%
  left_join(drive_dist) %>%
  select(n_SOV_HBO, n_HOV_HBO,
         n_SOV_HBW, n_HOV_HBW,
         n_SOV_NHB, n_HOV_NHB, miles) %>%
  mutate(PMT_SOV_HBO = n_SOV_HBO * miles,
         PMT_HOV_HBO = n_HOV_HBO * miles,
         PMT_SOV_HBW = n_SOV_HBW * miles,
         PMT_HOV_HBW = n_HOV_HBW * miles,
         PMT_SOV_NHB = n_SOV_NHB * miles,
         PMT_HOV_NHB = n_HOV_NHB * miles)%>%
  # filter out NA values because those rows have 0 SOV and HOV trips
  filter(is.na(PMT_HOV_HBO) != TRUE & is.na(PMT_SOV_HBO) != TRUE) %>%
  filter(is.na(PMT_HOV_HBW) != TRUE & is.na(PMT_SOV_HBW) != TRUE) %>%
  filter(is.na(PMT_HOV_NHB) != TRUE & is.na(PMT_SOV_NHB) != TRUE) %>%
  summarize(total_SOV_miles_HBO = sum(PMT_SOV_HBO),
            total_HOV_miles_HBO = sum(PMT_HOV_HBO),
            total_SOV_miles_HBW = sum(PMT_SOV_HBW),
            total_HOV_miles_HBW = sum(PMT_HOV_HBW),
            total_SOV_miles_NHB = sum(PMT_SOV_NHB),
            total_HOV_miles_NHB = sum(PMT_HOV_NHB))

PMT_car
```

##calculating existing VMT
```{r}
#using parameter for HBO, HBW and NHB for carpool- 2 plus only (daily trips) in NCHRP 716, Table 4-16
VMT_HBO <- PMT_car$total_SOV_miles_HBO[1] + 
       PMT_car$total_HOV_miles_HBO[1] / 2.71
VMT_HBW <- PMT_car$total_SOV_miles_HBW[1] + 
       PMT_car$total_HOV_miles_HBW[1] / 2.42
VMT_NHB <- PMT_car$total_SOV_miles_NHB[1] + 
       PMT_car$total_HOV_miles_NHB[1] / 1.66

VMT_sum <- VMT_HBO + VMT_HBW + VMT_NHB

VMT_HBO
VMT_HBW
VMT_NHB
VMT_sum
```


##Bike - existing
```{r}
PMT_bike <- skims %>%
  left_join(bike_dist, by = c("fromId", "toId")) %>%
  select(n_bike_HBO, n_bike_HBW, n_bike_NHB, miles) %>%
  mutate(PMT_bike_HBO = n_bike_HBO * miles,
         PMT_bike_HBW = n_bike_HBW * miles,
         PMT_bike_NHB = n_bike_NHB * miles)%>%
  # filter out NA values because those rows have 0 bike trips
  filter(is.na(PMT_bike_HBO) != TRUE) %>%
  filter(is.na(PMT_bike_HBW) != TRUE) %>%
  filter(is.na(PMT_bike_NHB) != TRUE) %>%
  summarize(total_bike_miles_HBO = sum(PMT_bike_HBO),
            total_bike_miles_HBW = sum(PMT_bike_HBW),
            total_bike_miles_NHB = sum(PMT_bike_NHB)
             )

PMT_bike
```

##Walk - existing
```{r}
PMT_walk <- skims %>%
  left_join(ped_dist, by = c("fromId", "toId")) %>%
  select(n_walk_HBO, n_walk_HBW, n_walk_NHB, miles) %>%
  mutate(PMT_walk_HBO = n_walk_HBO * miles,
         PMT_walk_HBW = n_walk_HBW * miles,
         PMT_walk_NHB = n_walk_NHB * miles)%>%
  filter(is.na(PMT_walk_HBO) != TRUE) %>%
  filter(is.na(PMT_walk_HBW) != TRUE) %>%
  filter(is.na(PMT_walk_NHB) != TRUE) %>%
  summarize(total_walk_miles_HBO = sum(PMT_walk_HBO),
            total_walk_miles_HBW = sum(PMT_walk_HBW),
            total_walk_miles_NHB = sum(PMT_walk_NHB))

PMT_walk
```

#Find VMT & PMt - Alternative

```{r}
# r5r_core_alternative <- here("alternative",
#                           "networks") %>%
#   setup_r5(verbose = FALSE)
# # 
# # ped_dist_alt <- detailed_itineraries(r5r_core_alternative,
# #                                  origins = origins,
# #                                  destinations = destinations,
# #                                  mode = "WALK",
# #                                  verbose = FALSE)  %>%
#   # mutate(miles = distance / 1609.34) %>%
#   # select(fromId, toId, miles) %>%
#   # st_drop_geometry()
# 
# # bike_dist_alt <- detailed_itineraries(r5r_core_alternative,
# #                                  origins = origins,
# #                                  destinations = destinations,
# #                                  mode = "BICYCLE",
# #                                  verbose = FALSE) %>%
# #   mutate(miles = distance / 1609.34) %>%
# #   select(fromId, toId, miles) %>%
# #   st_drop_geometry()
# 
# drive_dist_alt <- detailed_itineraries(r5r_core_alternative,
#                                  origins = origins,
#                                  destinations = destinations,
#                                  mode = "CAR",
#                                  verbose = FALSE) %>%
#   mutate(miles = distance / 1609.34) %>%
#   select(fromId, toId, miles) %>%
#   st_drop_geometry()
# 
# stop_r5() 
```

```{r}
# write_csv(drive_dist_alt, here("alternative",
#               "data",
#               "drive_dist_alt.csv"))
# 
# write_csv(ped_dist_alt, here("alternative",
#               "data",
#               "ped_dist_alt.csv"))
# 
# write_csv(drive_dist_alt, here("alternative",
#               "data",
#               "drive_dist_alt.csv"))

```

```{r}
bike_dist_alt <- here("alternative",
                     "data",
                     "bike_dist_alt.csv") %>%
  read_csv(
    #show_col_types = FALSE
    ) %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId))

drive_dist_alt <- here("alternative",
                     "data",
                     "drive_dist_alt.csv") %>%
  read_csv(
    #show_col_types = FALSE
    )  %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId))

ped_dist_alt <- here("alternative",
                     "data",
                     "ped_dist_alt.csv") %>%
  read_csv(
    #show_col_types = FALSE
    ) %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId))
```



## Car - alternative

```{r}
PMT_car_alt <- skims_alt %>%
  left_join(drive_dist_alt, by = c("fromId", "toId")) %>%
  select(n_SOV_HBO, n_HOV_HBO, 
         n_SOV_HBW, n_HOV_HBW,
         n_SOV_NHB, n_HOV_NHB, miles) %>%
  mutate(PMT_SOV_HBO_alt = n_SOV_HBO * miles,
         PMT_HOV_HBO_alt = n_HOV_HBO * miles,
         PMT_SOV_HBW_alt = n_SOV_HBW * miles,
         PMT_HOV_HBW_alt = n_HOV_HBW * miles,
         PMT_SOV_NHB_alt = n_SOV_NHB * miles,
         PMT_HOV_NHB_alt = n_HOV_NHB * miles)%>%
 # filter out NA values because those rows have 0 SOV and HOV trips
  filter(is.na(PMT_HOV_HBO_alt) != TRUE & is.na(PMT_SOV_HBO_alt) != TRUE) %>%
  filter(is.na(PMT_HOV_HBW_alt) != TRUE & is.na(PMT_SOV_HBW_alt) != TRUE) %>%
  filter(is.na(PMT_HOV_NHB_alt) != TRUE & is.na(PMT_SOV_NHB_alt) != TRUE) %>%
  summarize(total_SOV_miles_HBO_alt = sum(PMT_SOV_HBO_alt),
            total_HOV_miles_HBO_alt = sum(PMT_HOV_HBO_alt),
            total_SOV_miles_HBW_alt = sum(PMT_SOV_HBW_alt),
            total_HOV_miles_HBW_alt = sum(PMT_HOV_HBW_alt),
            total_SOV_miles_NHB_alt = sum(PMT_SOV_NHB_alt),
            total_HOV_miles_NHB_alt = sum(PMT_HOV_NHB_alt))

PMT_car_alt
```

##calculating alternative VMT
```{r}
#using parameter for HBO pm in NCHRP 716, Table 4-16
VMT_HBO_alt <- PMT_car_alt$total_SOV_miles_HBO_alt[1] + 
       PMT_car_alt$total_HOV_miles_HBO_alt[1] / 2.71
VMT_HBW_alt <- PMT_car_alt$total_SOV_miles_HBW_alt[1] + 
       PMT_car_alt$total_HOV_miles_HBW_alt[1] / 2.42
VMT_NHB_alt <- PMT_car_alt$total_SOV_miles_NHB_alt[1] + 
       PMT_car_alt$total_HOV_miles_NHB_alt[1] / 1.66

VMT_HBO_alt
VMT_HBW_alt
VMT_NHB_alt

```

##Bike - alternative

```{r}
PMT_bike_alt <- skims_alt %>%
  left_join(bike_dist_alt, by = c("fromId", "toId")) %>%
  select(n_bike_HBO, n_bike_HBW, n_bike_NHB, miles) %>%
  mutate(PMT_bike_HBO_alt = n_bike_HBO * miles,
         PMT_bike_HBW_alt = n_bike_HBW * miles,
         PMT_bike_NHB_alt = n_bike_NHB * miles)%>%
    # filter out NA values because those rows have 0 bike trips
  filter(is.na(PMT_bike_HBO_alt) != TRUE) %>%
  filter(is.na(PMT_bike_HBW_alt) != TRUE) %>%
  filter(is.na(PMT_bike_NHB_alt) != TRUE) %>%
  summarize(total_bike_miles_HBO_alt = sum(PMT_bike_HBO_alt),
            total_bike_miles_HBW_alt = sum(PMT_bike_HBW_alt),
            total_bike_miles_NHB_alt = sum(PMT_bike_NHB_alt)
             )

PMT_bike_alt
```

##Walk - alternative
```{r}
PMT_walk_alt <- skims_alt %>%
  left_join(ped_dist_alt, by = c("fromId", "toId")) %>%
  select(n_walk_HBO, n_walk_HBW, n_walk_NHB, miles) %>%
  mutate(PMT_walk_HBO_alt = n_walk_HBO * miles,
         PMT_walk_HBW_alt = n_walk_HBW * miles,
         PMT_walk_NHB_alt = n_walk_NHB * miles)%>%
  filter(is.na(PMT_walk_HBO_alt) != TRUE) %>%
  filter(is.na(PMT_walk_HBW_alt) != TRUE) %>%
  filter(is.na(PMT_walk_NHB_alt) != TRUE) %>%
  summarize(total_walk_miles_HBO_alt = sum(PMT_walk_HBO_alt),
            total_walk_miles_HBW_alt = sum(PMT_walk_HBW_alt),
            total_walk_miles_NHB_alt = sum(PMT_walk_NHB_alt))

PMT_walk_alt
```

#Save files
```{r}

write_csv(skims, here("existing",
              "data",
              "skims.csv"))
write_csv(PMT_car, here("existing",
              "data",
              "PMT_car.csv"))
write_csv(PMT_bike, here("existing",
              "data",
              "PMT_bike.csv"))
write_csv(PMT_walk, here("existing",
              "data",
              "PMT_walk.csv"))

write_csv(skims_alt, here("alternative",
              "data",
              "skims_alt.csv"))
write_csv(PMT_car_alt, here("alternative",
              "data",
              "PMT_car_alt.csv"))
write_csv(PMT_bike_alt, here("alternative",
              "data",
              "PMT_bike_alt.csv"))
write_csv(PMT_walk_alt, here("alternative",
              "data",
              "PMT_walk_alt.csv"))
```


## Visuals 

```{r}
hist <- ggplot(route_trip_by_purpose, aes(x=ridership_HBO)) + 
  geom_histogram(color = "white", fill = 'lightblue')

hist +theme_bw(base_size = 11) +
  scale_y_continuous(name = "Frequency",
                     breaks = seq(0, 100, by =10)) +
  scale_x_continuous(name = "Home-based nonwork trip transit ridership by routes")

ggsave("images/transit_ridership_HBO.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```
```{r}
hist <- ggplot(route_trip_by_purpose, aes(x=ridership_HBW)) + 
  geom_histogram(color = "white", fill = 'lightblue')

hist +theme_bw(base_size = 11) +
  scale_y_continuous(name = "Frequency",
                     breaks = seq(0, 100, by =10)) +
  scale_x_continuous(name = "Home-based work trip transit ridership by routes")

ggsave("images/transit_ridership_HBW.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```
```{r}
hist <- ggplot(route_trip_by_purpose, aes(x=ridership_NHB)) + 
  geom_histogram(color = "white", fill = 'lightblue')

hist +theme_bw(base_size = 11) +
  scale_y_continuous(name = "Frequency",
                     breaks = seq(0, 100, by =10)) +
  scale_x_continuous(name = "Non Home-based trip transit ridership by routes")

ggsave("images/transit_ridership_NHB.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```
```{r}
hist <- ggplot(route_trip_by_purpose, aes(x=sum_ridership)) + 
  geom_histogram(color = "white", fill = 'lightblue')

hist +theme_bw(base_size = 11) +
  scale_y_continuous(name = "Frequency",
                     breaks = seq(0, 100, by =10)) +
  scale_x_continuous(name = "All trips transit ridership by routes")

ggsave("images/transit_ridership_sum.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```


```{r}
ridership_change 
```

```{r}

map_data_PMT_bike <- skims %>%
  left_join(bike_dist, by = c("fromId", "toId")) %>%
  select(fromId,toId,n_bike_HBO,
         n_bike_HBW, 
         n_bike_NHB, miles) %>%
  mutate(PMT_bike_HBO = n_bike_HBO * miles,
         PMT_bike_HBW = n_bike_HBW * miles,
         PMT_bike_NHB = n_bike_NHB * miles)%>%
  group_by(fromId) %>%
  summarise(PMT_bike_HBO=sum(PMT_bike_HBO),
            PMT_bike_HBW=sum(PMT_bike_HBW),
            PMT_bike_NHB=sum(PMT_bike_NHB),
            PMT_bike_all = sum(PMT_bike_HBO,PMT_bike_HBW,PMT_bike_NHB))%>%
  select(fromId,PMT_bike_HBO,PMT_bike_HBW,PMT_bike_NHB,PMT_bike_all ) %>%
 mutate(GEOID = as.character(fromId)) 
# filter(fromId != 49035110104 & fromId != 49035114800 & fromId != 49035100200)



zones <- here("zones",
              "boundaries.geojson") %>%
  st_read()%>%
inner_join(map_data_PMT_bike)
```
```{r}
map_palette <- brewer.pal(5, "PuBuGn")

map <- ggplot(zones) +
  geom_sf(aes(fill = PMT_bike_all),
          color = NA) +
  scale_fill_gradientn(colours = map_palette,
                       name = "PMT of bike mode \nfor all trips from tracts\n in existing scenario") +
  theme_void()

map
ggsave("images/PMT_bike_all_existing.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```
```{r}
map_data_PMT_walk <- skims %>%
  left_join(ped_dist, by = c("fromId", "toId")) %>%
  select(fromId,toId,n_walk_HBO,
         n_walk_HBW, 
         n_walk_NHB, miles) %>%
  mutate(PMT_ped_HBO = n_walk_HBO * miles,
         PMT_ped_HBW = n_walk_HBW * miles,
         PMT_ped_NHB = n_walk_NHB * miles)%>%
filter(is.na(PMT_ped_HBO)== FALSE |is.na(PMT_ped_HBW)== FALSE| is.na(PMT_ped_HBW)== FALSE)%>%
group_by(fromId) %>%
summarise(PMT_ped_HBO=sum(PMT_ped_HBO),
          PMT_ped_HBW=sum(PMT_ped_HBW),
          PMT_ped_NHB=sum(PMT_ped_NHB),
          PMT_ped_all = sum(PMT_ped_HBO,PMT_ped_HBW,PMT_ped_NHB))%>%
 select(fromId,PMT_ped_HBO,PMT_ped_HBW,PMT_ped_NHB,PMT_ped_all ) %>%
mutate(GEOID = as.character(fromId))




zones <- here("zones",
              "boundaries.geojson") %>%
  st_read()%>%
inner_join(map_data_PMT_walk)
```
```{r}
map_palette <- brewer.pal(5, "BrBG")

map <- ggplot(zones) +
  geom_sf(aes(fill = PMT_ped_all),
          color = NA) +
  scale_fill_gradientn(colours = map_palette,
                       name = "PMT of walk mode \nfor all trips from tracts\n in existing scenario") +
  theme_void()

map
ggsave("images/PMT_ped_all_existing.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```



```{r}
map_data_VMT_car <- skims %>%
  left_join(drive_dist, by = c("fromId", "toId")) %>%
  select(fromId,toId,n_SOV_HBO, n_HOV_HBO, 
         n_SOV_HBW, n_HOV_HBW,
         n_SOV_NHB, n_HOV_NHB, miles) %>%
  mutate(PMT_SOV_HBO = n_SOV_HBO * miles,
         PMT_HOV_HBO = n_HOV_HBO * miles,
         PMT_SOV_HBW = n_SOV_HBW * miles,
         PMT_HOV_HBW = n_HOV_HBW * miles,
         PMT_SOV_NHB = n_SOV_NHB * miles,
         PMT_HOV_NHB = n_HOV_NHB * miles)%>%
  mutate(VMT_HBO = PMT_SOV_HBO + PMT_HOV_HBO/2.71,
         VMT_HBW = PMT_SOV_HBW + PMT_HOV_HBW/2.42,
         VMT_NHB = PMT_SOV_NHB + PMT_HOV_NHB/1.66
         )%>%
  group_by(fromId) %>%
  summarise(VMT_HBO=sum(VMT_HBO),
          VMT_HBW=sum(VMT_HBW),
          VMT_NHB=sum(VMT_NHB),
          VMT_all = sum(VMT_HBO,VMT_HBW,VMT_NHB))%>%
  select(fromId,VMT_HBO,VMT_HBW,VMT_NHB,VMT_all) %>%
 mutate(GEOID = as.character(fromId),
        VMT_all = round(VMT_all))


zones <- here("zones",
              "boundaries.geojson") %>%
  st_read()%>%
inner_join(map_data_VMT_car)
```

```{r}
map_palette <- brewer.pal(5, "PuBuGn")
options(scipen=999)

map <- ggplot(zones) +
  geom_sf(aes(fill = VMT_all),
          color = NA) +
  scale_fill_gradientn(colours = map_palette,
                       name = "VMT of car mode \nfor all trips from tracts\n in existing scenario") +
  theme_void()

map
ggsave("images/VMT_car_all_existing.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```
## alternative
```{r}

map_data_PMT_bike <- skims_alt %>%
  left_join(bike_dist_alt, by = c("fromId", "toId")) %>%
  select(fromId,toId,n_bike_HBO,
         n_bike_HBW, 
         n_bike_NHB, miles) %>%
  mutate(PMT_bike_HBO = n_bike_HBO * miles,
         PMT_bike_HBW = n_bike_HBW * miles,
         PMT_bike_NHB = n_bike_NHB * miles)%>%
  group_by(fromId) %>%
  summarise(PMT_bike_HBO=sum(PMT_bike_HBO),
            PMT_bike_HBW=sum(PMT_bike_HBW),
            PMT_bike_NHB=sum(PMT_bike_NHB),
            PMT_bike_all = sum(PMT_bike_HBO,PMT_bike_HBW,PMT_bike_NHB))%>%
  select(fromId,PMT_bike_HBO,PMT_bike_HBW,PMT_bike_NHB,PMT_bike_all ) %>%
 mutate(GEOID = as.character(fromId)) 
# filter(fromId != 49035110104 & fromId != 49035114800 & fromId != 49035100200)



zones <- here("zones",
              "boundaries.geojson") %>%
  st_read()%>%
inner_join(map_data_PMT_bike)
```

```{r}
map_palette <- brewer.pal(5, "PuBuGn")

map <- ggplot(zones) +
  geom_sf(aes(fill = PMT_bike_all),
          color = NA) +
  scale_fill_gradientn(colours = map_palette,
                       name = "PMT of bike mode \nfor all trips from tracts\n in dist_alt scenario") +
  theme_void()

map
ggsave("images/PMT_bike_all_dist_alt.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```

```{r}
map_data_PMT_walk <- skims_alt %>%
  left_join(ped_dist_alt, by = c("fromId", "toId")) %>%
  select(fromId,toId,n_walk_HBO,
         n_walk_HBW, 
         n_walk_NHB, miles) %>%
  mutate(PMT_ped_HBO = n_walk_HBO * miles,
         PMT_ped_HBW = n_walk_HBW * miles,
         PMT_ped_NHB = n_walk_NHB * miles)%>%
filter(is.na(PMT_ped_HBO)== FALSE |is.na(PMT_ped_HBW)== FALSE| is.na(PMT_ped_HBW)== FALSE)%>%
group_by(fromId) %>%
summarise(PMT_ped_HBO=sum(PMT_ped_HBO),
          PMT_ped_HBW=sum(PMT_ped_HBW),
          PMT_ped_NHB=sum(PMT_ped_NHB),
          PMT_ped_all = sum(PMT_ped_HBO,PMT_ped_HBW,PMT_ped_NHB))%>%
 select(fromId,PMT_ped_HBO,PMT_ped_HBW,PMT_ped_NHB,PMT_ped_all ) %>%
mutate(GEOID = as.character(fromId))




zones <- here("zones",
              "boundaries.geojson") %>%
  st_read()%>%
inner_join(map_data_PMT_walk)
```

```{r}
map_palette <- brewer.pal(5, "PuBuGn")

map <- ggplot(zones) +
  geom_sf(aes(fill = PMT_ped_all),
          color = NA) +
  scale_fill_gradientn(colours = map_palette,
                       name = "PMT of walk mode \nfor all trips from tracts\n in dist_alt scenario") +
  theme_void()

map
ggsave("images/PMT_ped_all_dist_alt.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```

```{r}
map_data_VMT_car <- skims_alt %>%
  left_join(drive_dist_alt, by = c("fromId", "toId")) %>%
  select(fromId,toId,n_SOV_HBO, n_HOV_HBO, 
         n_SOV_HBW, n_HOV_HBW,
         n_SOV_NHB, n_HOV_NHB, miles) %>%
  mutate(PMT_SOV_HBO = n_SOV_HBO * miles,
         PMT_HOV_HBO = n_HOV_HBO * miles,
         PMT_SOV_HBW = n_SOV_HBW * miles,
         PMT_HOV_HBW = n_HOV_HBW * miles,
         PMT_SOV_NHB = n_SOV_NHB * miles,
         PMT_HOV_NHB = n_HOV_NHB * miles)%>%
  mutate(VMT_HBO = PMT_SOV_HBO + PMT_HOV_HBO/2.71,
         VMT_HBW = PMT_SOV_HBW + PMT_HOV_HBW/2.42,
         VMT_NHB = PMT_SOV_NHB + PMT_HOV_NHB/1.66
         )%>%
  group_by(fromId) %>%
  summarise(VMT_HBO=sum(VMT_HBO),
          VMT_HBW=sum(VMT_HBW),
          VMT_NHB=sum(VMT_NHB),
          VMT_all = sum(VMT_HBO,VMT_HBW,VMT_NHB))%>%
  select(fromId,VMT_HBO,VMT_HBW,VMT_NHB,VMT_all) %>%
 mutate(GEOID = as.character(fromId),
        VMT_all = round(VMT_all))


zones <- here("zones",
              "boundaries.geojson") %>%
  st_read()%>%
inner_join(map_data_VMT_car)
```

```{r}
map_palette <- brewer.pal(5, "PuBuGn")
options(scipen=999)

map <- ggplot(zones) +
  geom_sf(aes(fill = VMT_all),
          color = NA) +
  scale_fill_gradientn(colours = map_palette,
                       name = "VMT of car mode \nfor all trips from tracts\n in dist_alt scenario") +
  theme_void()

map
ggsave("images/VMT_car_all_dist_alt.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```

