---
title: "Assignment 5"
author: "Charuvi Begwani, Aijing Li"
date: "2/22/2022"
output: html_document
---

#Accessibility 
In this file, we calculate accessibility to jobs from the centroid of each traffic analysis zone for each mode in the existing and alternate scenario.


#Load relevant libraries

```{r}
options(java.parameters = '-Xmx2G')

library(tidyverse)
library(sf)
library(tigris)
library(osmdata)
library(devtools)
library(tidytransit)
library(here)
library(r5r)
library(lubridate)
library(scenRios)
library(knitr)
library(ggthemes)
library(RColorBrewer)

```

#Set up centroid information

```{r}
jobs <- here("existing",
             "data",
             "zone_data.csv") %>%
  read_csv() %>%
  select(GEOID, total_emp) %>%
  mutate(GEOID = as.character(GEOID))%>%
  mutate(id = GEOID)%>%
  select(id, total_emp)

centroids <- here("zones",
                  "centroids.geojson") %>%
  st_read() %>%
  left_join(jobs)

```

#Set up r5r core

```{r}
existing_core <- here("existing",
                      "networks") %>%
  setup_r5(verbose = FALSE)
```

#Calculate accessibility for cars - Existing

```{r}
car_access_existing <- accessibility(existing_core,
                                     origins = centroids,
                                     destinations = centroids,
                                     opportunities_colname = 'total_emp',
                                     mode = "CAR",
                                     decay_function = "logistic",
                                     cutoffs = 35,
                                     decay_value = 5, 
                                     verbose = FALSE)

```

#Calculate accessibility for transit - Existing

```{r}
transit_access_existing <- accessibility(existing_core,
                                     origins = centroids,
                                     destinations = centroids,
                                     opportunities_colname = 'total_emp',
                                     mode = "TRANSIT",
                                     decay_function = "logistic",
                                     cutoffs = 50,
                                     decay_value = 8, 
                                     verbose = FALSE,
                                     departure_datetime = 
                                       ymd_hm("2022-03-09 17:00"),
                                     time_window = 120)
```


```{r}
stop_r5()

```


#Set up r5r core - alternative

```{r}
alternative_core <- here("alternative",
                      "networks") %>%
  setup_r5(verbose = FALSE)
```
#Calculate accessibility for cars - Alternative

```{r}
car_access_alternative <- accessibility(alternative_core,
                                     origins = centroids,
                                     destinations = centroids,
                                     opportunities_colname = 'total_emp',
                                     mode = "CAR",
                                     decay_function = "logistic",
                                     cutoffs = 35,
                                     decay_value = 5, 
                                     verbose = FALSE)

```

#Calculate accessibility for transit - Alternative

```{r}
transit_access_alternative <- accessibility(alternative_core,
                                     origins = centroids,
                                     destinations = centroids,
                                     opportunities_colname = 'total_emp',
                                     mode = "TRANSIT",
                                     decay_function = "logistic",
                                     cutoffs = 50,
                                     decay_value = 8, 
                                     verbose = FALSE,
                                     departure_datetime = 
                                       ymd_hm("2022-03-09 17:00"),
                                     time_window = 120)
```

```{r}
stop_r5()

```


#Compare auto and transit accessibility - Existing

```{r}
car_access_existing <- car_access_existing %>%
  select(from_id, accessibility) %>%
  rename(GEOID = from_id,
         car_access = accessibility)

transit_access_existing <- transit_access_existing %>%
  select(from_id, accessibility) %>%
  rename(GEOID = from_id,
         transit_access = accessibility)

access_compare_ex <- left_join(car_access_existing, transit_access_existing) %>%
  mutate(ratio = transit_access / car_access)
```

#Compare auto and transit accessibility - Alternative

```{r}
car_access_alternative <- car_access_alternative %>%
  select(from_id, accessibility) %>%
  rename(GEOID = from_id,
         car_access = accessibility)

transit_access_alternative <- transit_access_alternative %>%
  select(from_id, accessibility) %>%
  rename(GEOID = from_id,
         transit_access = accessibility)

access_compare_alt <- left_join(car_access_alternative, transit_access_alternative) %>%
  mutate(ratio = transit_access / car_access)
```

#summarize existing accessibility

```{r}

```

#summarize alternative accessibility

```{r}

```

#Chloropleth map showing spatial variation in car accessibility - existing

```{r}

```

#Chloropleth map showing spatial variation in car accessibility - alternative

```{r}

```

#Chloropleth map showing spatial variation in transit accessibility - existing

```{r}

```

#Chloropleth map showing spatial variation in transit accessibility - alternative

```{r}

```

#Chloropleth map on spatial variation in ratio of car and transit accessibility - existing

```{r}

```

#Chloropleth map on spatial variation in ratio of car and transit accessibility - alternative

```{r}

```



#Save your results
```{r}
write_csv(access_compare_ex, file = here("existing", 
                                      "data",
                                      "access_compare_ex.csv"))
write_csv(access_compare_alt, file = here("alternative", 
                                      "data",
                                      "access_compare_alt.csv"))
```

