---
title: "Assignment 10"
author: "Charuvi Begwani, Aijing Li"
date: "4/11/2022"
output: html_document
---

#Load libraries

```{r}
options(java.parameters = "-Xmx2G")

library(here)
library(tidyverse)
library(stplanr)
library(r5r)
library(sf)
library(tidyverse)
```

#Load data from previous assignments including production and attraction zone IDs, the transit routes included in that transit trip, and the number of trips by each mode for each trip purpose.

```{r}
skims <- here("existing",
                     "data",
                     "skims.csv") %>%
  read_csv(
    #show_col_types = FALSE
    ) %>%
  select(fromId, 
         toId, 
         n_transit_HBO,
         n_SOV_HBO,
         n_HOV_HBO,
         n_walk_HBO,
         n_bike_HBO,
         n_transit_HBW,
         n_SOV_HBW,
         n_HOV_HBW,
         n_walk_HBW,
         n_bike_HBW,
         n_transit_NHB,
         n_SOV_NHB,
         n_HOV_NHB,
         n_walk_NHB,
         n_bike_NHB,
         routes) %>%
  mutate(fromId = as.character(fromId),
         toId = as.character(toId))
```

#Convert P-A matrix to O-D matrix: HBO Trips

##Convert data frame to matrix
```{r}
HBO_PA_mat <- skims %>%
  od_to_odmatrix(attrib = "n_transit_HBO", 
                 name_orig = "fromId",
                 name_dest = "toId") 

HBO_PA_mat <- HBO_PA_mat[,row.names(HBO_PA_mat)]
```

##Transpose trip matrix
```{r}
HBO_PA_mat_trans <- t(HBO_PA_mat)
```

##Average matrix with its transpose 
```{r}
HBO_OD_mat <- (HBO_PA_mat + HBO_PA_mat_trans) / 2
```

##Convert matrix to data frame
```{r}
HBO_OD_table <- HBO_OD_mat %>%
  odmatrix_to_od() %>%
  rename(fromId = orig,
         toId = dest) %>%
  left_join(skims) %>%
  select(-n_transit_HBO)
```
#Convert P-A matrix to O-D matrix: HBW Trips

```{r}
##Convert data frame to matrix
HBW_PA_mat <- skims %>%
  od_to_odmatrix(attrib = "n_transit_HBW", 
                 name_orig = "fromId",
                 name_dest = "toId") 

HBW_PA_mat <- HBW_PA_mat[,row.names(HBW_PA_mat)]

##Transpose trip matrix
HBW_PA_mat_trans <- t(HBW_PA_mat)

##Average matrix with its transpose and convert matrix to data frame
HBW_OD_mat <- (HBW_PA_mat + HBW_PA_mat_trans) / 2

HBW_OD_table <- HBW_OD_mat %>%
  odmatrix_to_od() %>%
  rename(fromId = orig,
         toId = dest) %>%
  left_join(skims) %>%
  select(-n_transit_HBW)

```
#Convert P-A matrix to O-D matrix: NHB Trips

```{r}
##Convert data frame to matrix
NHB_PA_mat <- skims %>%
  od_to_odmatrix(attrib = "n_transit_NHB", 
                 name_orig = "fromId",
                 name_dest = "toId") 

NHB_PA_mat <- NHB_PA_mat[,row.names(NHB_PA_mat)]

##Transpose trip matrix
NHB_PA_mat_trans <- t(NHB_PA_mat)

##Average matrix with its transpose and convert matrix to data frame
NHB_OD_mat <- (NHB_PA_mat + NHB_PA_mat_trans) / 2

NHB_OD_table <- NHB_OD_mat %>%
  odmatrix_to_od() %>%
  rename(fromId = orig,
         toId = dest) %>%
  left_join(skims) %>%
  select(-n_transit_NHB)

```

#Count trips using each route
```{r}
route_trips_HBO <- HBO_OD_table %>%
  filter(flow > 0 & !is.na(routes)) %>%
  mutate(route_1 = str_split_fixed(routes, "\\|", 3)[,1],
         route_2 = str_split_fixed(routes, "\\|", 3)[,2],
         route_3 = str_split_fixed(routes, "\\|", 3)[,3]) %>%
  pivot_longer(cols = c(route_1, route_2, route_3),
               values_to = "route") %>%
  filter(route != "") %>%
  select(route, flow) %>%
  group_by(route) %>%
  summarize(ridership_HBO = round(sum(flow)))

route_trips_HBW <- HBW_OD_table %>%
  filter(flow > 0 & !is.na(routes)) %>%
  mutate(route_1 = str_split_fixed(routes, "\\|", 3)[,1],
         route_2 = str_split_fixed(routes, "\\|", 3)[,2],
         route_3 = str_split_fixed(routes, "\\|", 3)[,3]) %>%
  pivot_longer(cols = c(route_1, route_2, route_3),
               values_to = "route") %>%
  filter(route != "") %>%
  select(route, flow) %>%
  group_by(route) %>%
  summarize(ridership_HBW = round(sum(flow)))

route_trips_NHB <- NHB_OD_table %>%
  filter(flow > 0 & !is.na(routes)) %>%
  mutate(route_1 = str_split_fixed(routes, "\\|", 3)[,1],
         route_2 = str_split_fixed(routes, "\\|", 3)[,2],
         route_3 = str_split_fixed(routes, "\\|", 3)[,3]) %>%
  pivot_longer(cols = c(route_1, route_2, route_3),
               values_to = "route") %>%
  filter(route != "") %>%
  select(route, flow) %>%
  group_by(route) %>%
  summarize(ridership_NHB = round(sum(flow)))

```
## join all modes
```{r}
#join all table and sum 
route_trip_by_purporse<-left_join(route_trips_HBO,route_trips_HBW,by="route")%>%
                        left_join(route_trips_NHB,by="route")%>%
                        select(route,ridership_HBO,ridership_HBW,ridership_NHB)%>%
                        mutate(sum_ridership=ridership_HBO+ridership_HBW+ridership_NHB)
                        
```

```{r}
route_trip_by_purporse
```

#Find VMT and PMT by mode
##Calculate trip distances

```{r}
centroids <- here("zones",
                   "centroids.geojson") %>%
  st_read() %>%
  filter(!st_is_empty(.))

origins <- centroids %>%
  slice(rep(1:n(), each = n())) %>%
  mutate(dest_order =
           rep(seq(1:length(centroids$id)),
               length(centroids$id)))

destinations <- origins %>%
  arrange(dest_order)

r5r_core_existing <- here("existing",
                          "networks") %>%
  setup_r5(verbose = FALSE)

ped_dist <- detailed_itineraries(r5r_core_existing,
                                 origins = origins,
                                 destinations = destinations,
                                 mode = "WALK",
                                 verbose = FALSE,
                                 max_trip_duration = 150)  %>%
  mutate(miles = distance / 1609.34) %>%
  select(fromId, toId, miles) %>%
  st_drop_geometry()

drive_dist <- detailed_itineraries(r5r_core_existing,
                                 origins = origins,
                                 destinations = destinations,
                                 mode = "CAR",
                                 verbose = FALSE,
                                 max_trip_duration = 150) %>%
  mutate(miles = distance / 1609.34) %>%
  select(fromId, toId, miles) %>%
  st_drop_geometry()

bike_dist <- detailed_itineraries(r5r_core_existing,
                                 origins = origins,
                                 destinations = destinations,
                                 mode = "BICYCLE",
                                 verbose = FALSE,
                                 max_trip_duration = 150) %>%
  mutate(miles = distance / 1609.34) %>%
  select(fromId, toId, miles) %>%
  st_drop_geometry()

stop_r5() 
```



## Calculate PMT by mode

## HBO

```{r}
PMT_car_HBO <- skims %>%
  left_join(drive_dist, by = c("fromId", "toId")) %>%
  select(n_SOV_HBO, n_HOV_HBO, miles) %>%
  mutate(PMT_SOV_HBO = n_SOV_HBO * miles,
         PMT_HOV_HBO = n_HOV_HBO * miles)%>%
  # filter out NA values because those rows have 0 SOV and HOV trips
  filter(is.na(PMT_HOV_HBO) != TRUE & is.na(PMT_SOV_HBO) != TRUE) %>%
  summarize(total_SOV_miles = sum(PMT_SOV_HBO),
             total_HOV_miles = sum(PMT_HOV_HBO))


PMT_car_HBO
```

```{r}
#using parameter for HBO pm in NCHRP 716, Table 4-16
VMT_HBO <- PMT_car_HBO$total_SOV_miles[1] + 
       PMT_car_HBO$total_HOV_miles[1] / 1.66

VMT_HBO
```
```{r}
PMT_bike_HBO <- skims %>%
  left_join(drive_dist, by = c("fromId", "toId")) %>%
  select(n_bike_HBO, miles) %>%
  mutate(PMT_bike_HBO = n_bike_HBO * miles,
         )%>%
  # filter out NA values because those rows have 0 bike trips
  filter(is.na(PMT_bike_HBO) != TRUE) %>%
  summarize(total_bike_miles = sum(PMT_bike_HBO)
             )


PMT_bike_HBO
```

```{r}
PMT_walk_HBO <- skims %>%
  left_join(drive_dist, by = c("fromId", "toId")) %>%
  select(n_walk_HBO, miles) %>%
  mutate(PMT_walk_HBO = n_walk_HBO * miles,
         )%>%
  filter(is.na(PMT_walk_HBO) != TRUE) %>%
  summarize(total_walk_miles = sum(PMT_walk_HBO)
             )


PMT_walk_HBO
```

## HBW
```{r}
PMT_car_HBW <- skims %>%
  left_join(drive_dist, by = c("fromId", "toId")) %>%
  select(n_SOV_HBW, n_HOV_HBW, miles) %>%
  mutate(PMT_SOV_HBW = n_SOV_HBW * miles,
         PMT_HOV_HBW = n_HOV_HBW * miles)%>%
  # filter out NA values because those rows have 0 SOV and HOV trips
  filter(is.na(PMT_HOV_HBW) != TRUE & is.na(PMT_SOV_HBW) != TRUE) %>%
  summarize(total_SOV_miles = sum(PMT_SOV_HBW),
             total_HOV_miles = sum(PMT_HOV_HBW))


PMT_car_HBW

#using parameter for HBW pm for all auto modes in NCHRP 716, Table 4-16
VMT_HBW <- PMT_car_HBW$total_SOV_miles[1] + 
       PMT_car_HBW$total_HOV_miles[1] / 1.11

VMT_HBW
```
```{r}
PMT_bike_HBW <- skims %>%
  left_join(drive_dist, by = c("fromId", "toId")) %>%
  select(n_bike_HBW, miles) %>%
  mutate(PMT_bike_HBW = n_bike_HBW * miles,
         )%>%
  filter(is.na(PMT_bike_HBO) != TRUE) %>%
  summarize(total_bike_miles = sum(PMT_bike_HBO)
             )


PMT_bike_HBW
```

```{r}
PMT_walk_HBW <- skims %>%
  left_join(drive_dist, by = c("fromId", "toId")) %>%
  select(n_walk_HBW, miles) %>%
  mutate(PMT_walk_HBW = n_walk_HBW * miles,
         )%>%
  filter(is.na(PMT_walk_HBW) != TRUE) %>%
  summarize(total_walk_miles = sum(PMT_walk_HBW)
             )


PMT_walk_HBW
```

## NHB
```{r}
PMT_car_NHB <- skims %>%
  left_join(drive_dist, by = c("fromId", "toId")) %>%
  select(n_SOV_NHB, n_HOV_NHB, miles) %>%
  mutate(PMT_SOV_NHB = n_SOV_NHB * miles,
         PMT_HOV_NHB = n_HOV_NHB * miles)%>%
  # filter out NA values because those rows have 0 SOV and HOV trips
  filter(is.na(PMT_HOV_NHB) != TRUE & is.na(PMT_SOV_NHB) != TRUE) %>%
  summarize(total_SOV_miles = sum(PMT_SOV_NHB),
             total_HOV_miles = sum(PMT_HOV_NHB))


PMT_car_NHB

#using parameter for NHB pm in NCHRP 716, Table 4-16
VMT_NHB <- PMT_car_NHB$total_SOV_miles[1] + 
       PMT_car_NHB$total_HOV_miles[1] / 1.65

VMT_NHB
```
```{r}
PMT_bike_NHB <- skims %>%
  left_join(drive_dist, by = c("fromId", "toId")) %>%
  select(n_bike_NHB, miles) %>%
  mutate(PMT_bike_NHB = n_bike_NHB * miles,
         )%>%
  filter(is.na(PMT_bike_HBO) != TRUE) %>%
  summarize(total_bike_miles = sum(PMT_bike_HBO)
             )


PMT_bike_NHB
```

```{r}
PMT_walk_NHB <- skims %>%
  left_join(drive_dist, by = c("fromId", "toId")) %>%
  select(n_walk_NHB, miles) %>%
  mutate(PMT_walk_NHB = n_walk_NHB * miles,
         )%>%
  filter(is.na(PMT_walk_NHB) != TRUE) %>%
  summarize(total_walk_miles = sum(PMT_walk_NHB)
             )


PMT_walk_NHB
```


```{r}
PMT_car <- rbind(PMT_car_HBO,PMT_car_HBW,PMT_car_NHB)
PMT_bike <-rbind(PMT_bike_HBO,PMT_bike_HBW,PMT_bike_NHB)
PMT_walk <-rbind(PMT_walk_HBO,PMT_walk_HBW,PMT_walk_NHB)
#using parameter for all trips for all auto modes in NCHRP 716, Table 4-16
VMT_all <- PMT_car$total_SOV_miles[1] + 
       PMT_car$total_HOV_miles[1] / 1.5
PMT_car
PMT_bike
PMT_walk
```

```{r}

write_csv(skims, here("existing",
              "data",
              "skims.csv"))
write_csv(PMT_car, here("existing",
              "data",
              "PMT_car.csv"))
write_csv(PMT_bike, here("existing",
              "data",
              "PMT_bike.csv"))
write_csv(PMT_walk, here("existing",
              "data",
              "PMT_walk.csv"))
```

```{r}
hist <- ggplot(route_trip_by_purporse, aes(x=ridership_HBO)) + 
  geom_histogram(color = "white", fill = 'lightblue')

hist +theme_bw(base_size = 11) +
  scale_y_continuous(name = "Frequency",
                     breaks = seq(0, 100, by =10)) +
  scale_x_continuous(name = "Home-based nonwork trip transit ridership by routes")

ggsave("images/transit_ridership_HBO.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```
```{r}
hist <- ggplot(route_trip_by_purporse, aes(x=ridership_HBW)) + 
  geom_histogram(color = "white", fill = 'lightblue')

hist +theme_bw(base_size = 11) +
  scale_y_continuous(name = "Frequency",
                     breaks = seq(0, 100, by =10)) +
  scale_x_continuous(name = "Home-based work trip transit ridership by routes")

ggsave("images/transit_ridership_HBW.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```
```{r}
hist <- ggplot(route_trip_by_purporse, aes(x=ridership_NHB)) + 
  geom_histogram(color = "white", fill = 'lightblue')

hist +theme_bw(base_size = 11) +
  scale_y_continuous(name = "Frequency",
                     breaks = seq(0, 100, by =10)) +
  scale_x_continuous(name = "Non Home-based trip transit ridership by routes")

ggsave("images/transit_ridership_NHB.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```
```{r}
hist <- ggplot(route_trip_by_purporse, aes(x=sum_ridership)) + 
  geom_histogram(color = "white", fill = 'lightblue')

hist +theme_bw(base_size = 11) +
  scale_y_continuous(name = "Frequency",
                     breaks = seq(0, 100, by =10)) +
  scale_x_continuous(name = "All trips transit ridership by routes")

ggsave("images/transit_ridership_sum.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```


