---
title: "Assignment 7"
author: "Aijing Li, Charuvi Begwani"
date: "3/18/2022"
output: html_document
---
```{r, message=FALSE,warning=FALSE}
library(here)
library(tidyverse)
library(survey)
library(srvyr)
library(naniar)
library(jtools)
library(knitr)
library(sf)
library(RColorBrewer)
```

# Trip productions
## Load NHTS data
```{r, message= FALSE, warning=FALSE, echo=FALSE}
trips <- read_csv(here("existing/data/trippub.csv"))
hhs <- read_csv(here("existing/data/hhpub.csv"))

```

## Select household-level variables
```{r}
saltlake_hhs <- hhs %>%
  filter(HH_CBSA == "41620") %>%
  mutate(inc_quint_ = case_when(HHFAMINC == "01" ~ "1st",
                               HHFAMINC == "02" ~ "1st",
                               HHFAMINC == "03" ~ "1st",
                               HHFAMINC == "04" ~ "1st",
                               HHFAMINC == "05" ~ "2nd",
                               HHFAMINC == "06" ~ "2nd",
                               HHFAMINC == "07" ~ "3rd",
                               HHFAMINC == "08" ~ "4th",
                               HHFAMINC == "09" ~ "4th",
                               HHFAMINC == "10" ~ "5th",
                               HHFAMINC == "11" ~ "5th",
                               TRUE ~ "NA")) %>%
  mutate(size_ = case_when(HHSIZE == 1 ~ "one",
                              HHSIZE == 2 ~ "two",
                              HHSIZE == 3 ~ "three",
                              TRUE ~ "four_plus")) %>%
  mutate(zero_veh_ = (HHVEHCNT == 0)) %>%
  replace_with_na(list(inc_quint_ = "NA")) %>%
  select(HOUSEID, zero_veh_, size_, inc_quint_, WTHHFIN) 
```

## Select trip-level variables
```{r}
trips_by_purpose <- trips %>% 
  filter(HH_CBSA == "41620") %>%
  select(HOUSEID, WHYFROM, WHYTO) %>%
  mutate(home_based = case_when(WHYTO == "01" ~ TRUE,
                                WHYTO == "02" ~ TRUE,
                                WHYFROM == "01" ~ TRUE,
                                WHYFROM == "02" ~ TRUE,
                                TRUE ~ FALSE)) %>%
  mutate(work = ifelse(WHYTO == "03" | WHYFROM == "03", TRUE, FALSE)) %>%
  mutate(purpose = case_when(home_based & work ~ "HBW",
                            home_based ~ "HBO",
                            TRUE ~ "NHB"))%>%
  group_by(HOUSEID, purpose) %>%
  summarize(n = n()) %>%
  pivot_wider(names_from = "purpose", values_from = "n") 
```


```{r}
 hh_trips <- left_join(saltlake_hhs, trips_by_purpose) %>%
  replace_na(list(HBW = 0,
                  HBO = 0,
                  NHB = 0))
```
## Create a survey object
```{r}
svy_trips <- hh_trips %>%
  as_survey(weights = WTHHFIN)
```

## Estimate a household-level regression model
## for HBO trips
```{r}

HBO_model1 <- svyglm(HBO ~ zero_veh_ + size_ + inc_quint_, svy_trips)

export_summs(HBO_model1, 
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = "Full model")

```
```{r}
HBO_model2 <- svyglm(HBO ~ size_, svy_trips)

export_summs(HBO_model1, HBO_model2,
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = c("Full model", "Reduced model"))
```

## for HBW
```{r}

HBW_model1 <- svyglm(HBW ~ zero_veh_ + size_ + inc_quint_, svy_trips)

export_summs(HBW_model1, 
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = "Full model")
```
```{r}
HBW_model2 <- svyglm(HBW ~ size_ + inc_quint_, svy_trips)

export_summs(HBW_model1, HBW_model2,
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = c("Full model", "Reduced model"))
```
```{r}
HBW_model3 <- svyglm(HBW ~ 1, svy_trips)


export_summs(HBW_model1, HBW_model2, HBW_model3,
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = c("Full model", "Reduced model","Empty Model"))
```


## for NHB
```{r}
NHB_model1 <- svyglm(NHB ~ zero_veh_ + size_ + inc_quint_, svy_trips)

export_summs(NHB_model1, 
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = "Full model")

```
```{r}
NHB_model2 <- svyglm(NHB ~ size_+inc_quint_, svy_trips)

export_summs(NHB_model1, NHB_model2,
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = c("Full model", "Reduced model"))
```
`
```{r}
existing_zones <- here("existing",
                       "data",
                       "zone_data.csv") %>%
  read_csv() %>%
  mutate(hbo_prod = total_hhsE * HBO_model2$coefficients["(Intercept)"] +
                    hh_1personE * HBO_model2$coefficients["size_one"] +
                    hh_2personE * HBO_model2$coefficients["size_two"] +
                    hh_3personE * HBO_model2$coefficients["size_three"]
                    )%>%
  mutate(hbw_prod = total_hhsE * HBW_model1$coefficients["(Intercept)"] +
                    hh_1personE * HBW_model1$coefficients["size_one"] +
                    hh_2personE * HBW_model1$coefficients["size_two"] +
                    hh_3personE * HBW_model1$coefficients["size_three"]+
                    inc_quint_2 * HBW_model1$coefficients["inc_quint_2nd"]+
                    inc_quint_3 * HBW_model1$coefficients["inc_quint_3rd"]+
                    inc_quint_4 * HBW_model1$coefficients["inc_quint_4th"]+
                    inc_quint_5 * HBW_model1$coefficients["inc_quint_5th"]
           )%>%
    mutate(nhb_prod = total_hhsE * NHB_model2$coefficients["(Intercept)"] +
                    hh_1personE * NHB_model2$coefficients["size_one"] +
                    hh_2personE * NHB_model2$coefficients["size_two"] +
                    hh_3personE * NHB_model2$coefficients["size_three"]+
                    inc_quint_2 * NHB_model2$coefficients["inc_quint_2nd"]+
                    inc_quint_3 * NHB_model2$coefficients["inc_quint_3rd"]+
                    inc_quint_4 * NHB_model2$coefficients["inc_quint_4th"]+
                    inc_quint_5 * NHB_model2$coefficients["inc_quint_5th"]
           )
```

# Trip Attractions
## We apply the attraction model from Table 4-4 of the National Cooperative Highway Research Program (NCHRP) Report 716: Travel Demand Forecasting, to predict the the number of trips attracted to each zone for each of three trip purposes. We estimate the number of trips using the models found under "All person trips", since we would be including non-motorized modes in our model. For Home-based work trips we use the one model that has been provided. To estimate HBO trip attraction, we use “Model 3” under Home-based Nonwork trips, since we don’t have school enrollment data. For NHB trips, we use Model 1 under Non-home based trips, since we have data on basic employment.
```{r}
existing_zones <- existing_zones %>%
  replace_na(list(basic_emp = 0,
                  retail_emp = 0, 
                  service_emp = 0)) %>%
  mutate(hbo_attr = 0.7 * total_hhsE +
                    0.7 * basic_emp +
                    8.4 * retail_emp +
                    3.5 * service_emp) %>%
  mutate(hbw_attr = 1.2 * (basic_emp + 
                             retail_emp + 
                             service_emp)) %>%
  mutate(nhb_attr = 0.6 * total_hhsE +
                    0.5 * basic_emp +
                    4.7 * retail_emp +
                    1.4 * service_emp)
```


# Balancing productions and attractions
```{r}
trip_end_summary <- tibble(Purpose = c("HBO"),
                           Productions = c(sum(existing_zones$hbo_prod)),
                           Attractions = c(sum(existing_zones$hbo_attr)),
                           Difference = c(sum(existing_zones$hbo_attr) - 
                                            sum(existing_zones$hbo_prod)))

kable(trip_end_summary, format.args = list(big.mark = ","))
```

```{r}
trip_end_summary <- tibble(Purpose = c("HBW"),
                           Productions = c(sum(existing_zones$hbw_prod)),
                           Attractions = c(sum(existing_zones$hbw_attr)),
                           Difference = c(sum(existing_zones$hbw_attr) - 
                                            sum(existing_zones$hbw_prod)))
kable(trip_end_summary, format.args = list(big.mark = ","))
```

```{r}
trip_end_summary <- tibble(Purpose = c("NHB"),
                           Productions = c(sum(existing_zones$nhb_prod)),
                           Attractions = c(sum(existing_zones$nhb_attr)),
                           Difference = c(sum(existing_zones$nhb_attr) - 
                                            sum(existing_zones$nhb_prod)))
kable(trip_end_summary, format.args = list(big.mark = ","))

```
## scale trip attraction
```{r}
existing_zones <- existing_zones %>%
  mutate(hbo_attr_bal = hbo_attr * 
           sum(existing_zones$hbo_prod) / sum(existing_zones$hbo_attr)) %>%
  mutate(hbw_attr_bal = hbw_attr * 
           sum(existing_zones$hbw_prod) / sum(existing_zones$hbw_attr)) %>%
  mutate(nhb_attr_bal = nhb_attr * 
           sum(existing_zones$nhb_prod) / sum(existing_zones$nhb_attr)) 

```

##Check scaled totals

```{r}
trip_end_summary <- tibble(Purpose = c("HBO"),
                           Productions = c(sum(existing_zones$hbo_prod)),
                           Attractions = c(sum(existing_zones$hbo_attr_bal)),
                           Difference = c(sum(existing_zones$hbo_attr_bal) - 
                                            sum(existing_zones$hbo_prod)))

kable(trip_end_summary, format.args = list(big.mark = ","))
```



```{r}
hist <- ggplot(existing_zones, aes(x=hbw_attr_bal)) + 
  geom_histogram(color = "white", fill = 'lightblue')

hist +theme_bw(base_size = 11) +
  scale_y_continuous(name = "Number of tracts",
                     breaks = seq(0, 100, by =10)) +
  scale_x_continuous(name = "Home-based work trip attraction(Balanced)")

ggsave("images/hbw_attr_bal.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```

```{r}
hist <- ggplot(existing_zones, aes(x=hbo_attr_bal)) + 
  geom_histogram(color = "white", fill = 'lightblue')

hist +theme_bw(base_size = 11) +
  scale_y_continuous(name = "Number of tracts",
                     breaks = seq(0, 100, by =10)) +
  scale_x_continuous(name = "Home-based non-work trip attraction(Balanced)",
                      )

ggsave("images/hbo_attr_bal.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```
```{r}
hist <- ggplot(existing_zones, aes(x=nhb_attr_bal)) + 
  geom_histogram(color = "white", fill = 'lightblue')

hist +theme_bw(base_size = 11) +
  scale_y_continuous(name = "Number of tracts",
                     breaks = seq(0, 100, by =10)) +
  scale_x_continuous(name = "Non home-based trip attraction(Balanced)"
                     )

ggsave("images/nhb_attr_bal.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```
```{r}
hist <- ggplot(existing_zones, aes(x=hbo_prod)) + 
  geom_histogram(color = "white", fill = 'lightblue')

hist +theme_bw(base_size = 11) +
  scale_y_continuous(name = "Number of tracts",
                     breaks = seq(0, 100, by =10)) +
  scale_x_continuous(name = "Home-based non-work trip production"
                     )

ggsave("images/hbo_prod.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```
```{r}
hist <- ggplot(existing_zones, aes(x=hbw_prod)) + 
  geom_histogram(color = "white", fill = 'lightblue')

hist +theme_bw(base_size = 11) +
  scale_y_continuous(name = "Number of tracts",
                     breaks = seq(0, 100, by =10)) +
  scale_x_continuous(name = "Home-based work trip production"
                     )

ggsave("images/hbw_prod.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```
```{r}
hist <- ggplot(existing_zones, aes(x=nhb_prod)) + 
  geom_histogram(color = "white", fill = 'lightblue')

hist +theme_bw(base_size = 11) +
  scale_y_continuous(name = "Number of tracts",
                     breaks = seq(0, 100, by =10)) +
  scale_x_continuous(name = "Non home-based trip production"
                     )

ggsave("images/nhb_prod.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```
```{r}
existing_zones <- existing_zones %>%
  mutate(GEOID = as.character(GEOID))

zones <- here("zones",
              "boundaries.geojson") %>%
  st_read() %>%
  inner_join(existing_zones)
```


```{r}
map_palette <- brewer.pal(5, "PuBuGn")

map_diff_model2 <- ggplot(zones) +
  geom_sf(aes(fill = hbw_attr_bal),
          color = NA) +
  scale_fill_gradientn(colours = map_palette,
                       name = "Home-based work trip \nattraction (balanced)") +
  theme_void()

map_diff_model2
ggsave("images/map_hbw_attr_bal.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```
```{r}
map_palette <- brewer.pal(5, "PuBuGn")

map <- ggplot(zones) +
  geom_sf(aes(fill = hbo_attr_bal),
          color = NA) +
  scale_fill_gradientn(colours = map_palette,
                       name = "Home-based \nnon-work trip \nattraction (balanced)") +
  theme_void()

map
ggsave("images/map_hbo_attr_bal.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```

```{r}
map_palette <- brewer.pal(5, "PuBuGn")

map <- ggplot(zones) +
  geom_sf(aes(fill = nhb_attr_bal),
          color = NA) +
  scale_fill_gradientn(colours = map_palette,
                       name = "Non home-based trip \nattraction (balanced)") +
  theme_void()

map
ggsave("images/map_nhb_attr_bal.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```
```{r}
map_palette <- brewer.pal(5, "PuBuGn")

map <- ggplot(zones) +
  geom_sf(aes(fill = nhb_prod),
          color = NA) +
  scale_fill_gradientn(colours = map_palette,
                       name = "Non home-based trip \nproduction") +
  theme_void()

map
ggsave("images/map_nhb_prod.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```

```{r}
map_palette <- brewer.pal(5, "PuBuGn")

map <- ggplot(zones) +
  geom_sf(aes(fill = hbo_prod),
          color = NA) +
  scale_fill_gradientn(colours = map_palette,
                       name = "Home-based \nnon-work trip \nproduction") +
  theme_void()

map
ggsave("images/map_hbo_prod.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```
```{r}
map_palette <- brewer.pal(5, "PuBuGn")

map <- ggplot(zones) +
  geom_sf(aes(fill = hbw_prod),
          color = NA) +
  scale_fill_gradientn(colours = map_palette,
                       name = "Home-based \nwork trip \nproduction") +
  theme_void()

map
ggsave("images/map_hbw_prod.png",
       width = 20, height = 15, units = "cm",dpi = 300)
```

