---
title: "Assignment 8"
author: "Aijing Li, Charuvi Begwani"
date: "3/24/2022"
output: html_document
---

## Set-up and Load Data
```{r, message=FALSE,warning=FALSE, echo=FALSE}
library(here)
library(tidyverse)
library(sf)
library(survey)
library(srvyr)
library(od)
library(ggspatial)
library(scenRios)
```

```{r}
zones <- here("existing",
              "data",
              "zone_data.csv") %>%
  read_csv()

skims <- here("alternative",
              "data",
              "skims.csv") %>%
  read_csv()
```
```{r}
trips <- read_csv(here("existing","data", "trippub.csv"))%>%
  filter(HH_CBSA == "41620")

trips <- trips %>%
  mutate(home_based = case_when(WHYTO == "01" ~ TRUE,
                                WHYTO == "02" ~ TRUE,
                                WHYFROM == "01" ~ TRUE,
                                WHYFROM == "02" ~ TRUE,
                                TRUE ~ FALSE)) %>%
  mutate(work = ifelse(WHYTO == "03" | WHYFROM == "03", TRUE, FALSE)) %>%
  mutate(purpose = case_when(home_based & work ~ "HBW",
                            home_based ~ "HBO",
                            TRUE ~ "NHB"))
```
```{r}
trips_svy <- trips %>%
  as_survey(weights = WTTRDFIN)

ttime_by_purpose <- trips_svy %>%
  group_by(purpose) %>%
  summarise(avg_time = survey_mean(TRVLCMIN))

ttime_by_purpose
```

## Change 0 car time to 1 min for PA pairs between different ID
```{r}
skims$car_time[skims$fromId != skims$toId &  skims$car_time == 0] <- 1
```

## Calculate minimum travel time between PA paris
```{r}
skims <- skims %>%
  mutate(min_time = pmin(transit_time, 
                         car_time,
                         bike_time,
                         walk_time,
                         na.rm = TRUE)) 
```


# Calculate friction factors
## Exponential function
```{r}
# m_HBO <- ttime_by_purpose$avg_time[ttime_by_purpose$purpose == "HBO"]
# 
# skims <- skims %>%
#   mutate(F_HBO = exp(-1 * m_HBO * min_time)) 
```
## Power function
```{r}
# skims <- skims %>%
#   mutate(F_NHB = min_time^-2)
```
## Gemma function
## We choose Large MPO 1 parameters for HBO & NHB and MPO 3 for HBW based on comparison of avergae travel time
```{r}
skims <- skims %>%
  mutate(F_HBO = min_time^-3.993*exp(-0.019*min_time)) 
skims <- skims %>%
  mutate(F_HBW = min_time^-0.156*exp(-0.045*min_time)) 
skims <- skims %>%
  mutate(F_NHB = min_time^-3.345*exp(-0.003*min_time)) 
```
## Estimate travel flows
```{r}
HBO_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "hbo_prod",
                            zone_d = "hbo_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBO",
                            tolerance = 0.01,
                            max_iter = 50000)

HBW_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "hbw_prod",
                            zone_d = "hbw_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBW",
                            tolerance = 0.01,
                            max_iter = 50000)

NHB_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "nhb_prod",
                            zone_d = "nhb_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_NHB",
                            tolerance = 0.01,
                            max_iter = 50000)
```
## Print out results
```{r}
tail(HBO_dist$convergence)
```

```{r}
tail(HBW_dist$convergence)
```
```{r}
tail(NHB_dist$convergence)
```
```{r}
head(HBO_dist$flows)
```
```{r}
table(HBO_dist$flows$flow > 0)
table(HBW_dist$flows$flow > 0)
table(NHB_dist$flows$flow > 0)
```


## Compare calculated travel times to observed travel times
## Averages from NHTS data
```{r}
ttime_by_purpose
```
```{r}
HBO_flows <- HBO_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         HBO_flow = flow)

skims <- skims %>%
  left_join(HBO_flows) %>%
  replace_na(list(HBO_flow = 0)) %>%
  mutate(total_time = HBO_flow * min_time)
```
## Averages from gravity model
```{r}
HBO_mean_time <- sum(skims$total_time) / sum(skims$HBO_flow)

HBO_mean_time
```
```{r}
HBW_flows <- HBW_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         HBW_flow = flow)

skims <- skims %>%
  left_join(HBW_flows) %>%
  replace_na(list(HBW_flow = 0)) %>%
  mutate(total_time = HBW_flow * min_time)

HBW_mean_time <- sum(skims$total_time) / sum(skims$HBW_flow)

HBW_mean_time 
```
```{r}
NHB_flows <- NHB_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         NHB_flow = flow)

skims <- skims %>%
  left_join(NHB_flows) %>%
  replace_na(list(NHB_flow = 0)) %>%
  mutate(total_time = NHB_flow * min_time)

NHB_mean_time <- sum(skims$total_time) / sum(skims$NHB_flow)

NHB_mean_time 
```

#Save skim csv files
```{r}
write_csv(skims, file = here("alternative", 
                             "data",
                             "skims.csv"))
```

