---
title: "Assignment 5"
author: "Charuvi Begwani, Aijing Li"
date: "2/22/2022"
output: html_document
---

#Accessibility 
In this file, we calculate accessibility to jobs from the centroid of each traffic analysis zone for each mode in the existing and alternate scenario.


#Load relevant libraries

```{r}
options(java.parameters = '-Xmx2G')

library(r5r)
library(here)
library(tidyverse)
library(sf)
library(lubridate)
```

#Set up centroid information

```{r}
jobs <- here("existing",
             "data",
             "zone_data.csv") %>%
  read_csv() %>%
  select(GEOID, total_emp) %>%
  mutate(GEOID = as.character(GEOID))

centroids <- here("zones",
                  "centroids.geojson") %>%
  st_read() %>%
  left_join(jobs) %>%
  rename(id = GEOID)
```

#Set up r5r core

```{r}
existing_core <- here("existing",
                      "networks") %>%
  setup_r5(verbose = FALSE)
```

#Calculate accessibility for cars

```{r}

```

#Calculate accessibility for transit

```{r}

```

#Compare auto and transit accessibility
```{r}

```

#Save your results
```{r}

```

