---
title: "Assignment 2"
author: "Aijing Li, Charuvi Begwani"
date: "2/4/2022"
output: html_document
---


```{r}
# install.packages("ggthemes")
# install.packages("RColorBrewer")
# install.packages("leaflet")
# install.packages("htmlwidgets")
# install.packages("tibble")
# install.packages("treemapify")
# install.packages("lwgeom")
```

```{r}
library(tibble)
library(tidycensus)
library(sf)
library(RColorBrewer)
library(ggthemes)
library(dplyr)
library(treemapify)
library(tidyverse)
library(knitr)

```



# Introduction 
Main Street is one of the key commercial streets in Salt Lake City with banks, major retailers and high pedestrian footfall. In a few weeks in September and October 2020, a block on the Main Street was shut down for cars on the weekends (Thursday - Saturday) from 6-10 p.m. Referred to as the Downtown SLC Open Streets program, this was done in an attempt to revitalize the downtown area and encourage patrons. The new spaces created, and outdoor dining and retail services initiated as a result of social distancing requirements due to the pandemic, were leveraged to “expand options for pedestrians” and benefit on-street businesses. In response to the success of this initiative, it was repeated the next year in April with the concerned stretch expanded to include South Temple to 400 South. As before, this was an attempt to invite pedestrians with some ongoing activities and support local businesses. This news article reports that Downtown Alliance, the group leading this, is considering asking city to make it a permanent, year-round “pedestrian zone.” 

We are interested in modelling Main Street pedestrianization and closure for cars to explore how impactful the plan for a permanent pedestrian zone may be. We think this is especially relevant since Open Streets have been implemented all across the country, post-pandemic, and many other cities are also considering similar pedestrianization projects with the success of outdoor dining and retail activities.

#Getting census variable names
We first decide to retrieve census data on population characteristics 
through the American Commmunity Survey 2019, using the tidycensus package.

```{r}
area_vars_2019 <- load_variables(2019, "acs5")
```
#Download census data.
We then download census data for the Salt Lake Metropolitan Statistical Area 
for selected variables that are relevant to this analysis, 
including household size, vehicle ownership and household income. 
This data is available at the census block level, and we retrieve it for 
Salt Lake and Tooele Counties that constitute the Salt Lake MSA.

```{r}

vars <- c(total_hhs = 'B08203_001',
          no_veh = 'B08203_002',
          avg_hh_size = "B25010_001",
          
          hh_1person = 'B08201_007',
          hh_2person = 'B08201_013',
          hh_3person = 'B08201_019',
          hh_4person_plus = 'B08201_025',
          
          inc_lt_10k = 'B19001_002',
          inc_btw_10k_15k = 'B19001_003',
          inc_btw_15k_20k = 'B19001_004',
          inc_btw_20k_25k = 'B19001_005',
          inc_btw_25k_30k = 'B19001_006',
          inc_btw_30k_35k = 'B19001_007',
          inc_btw_35k_40k = 'B19001_008',
          inc_btw_40k_45k = 'B19001_009',
          inc_btw_45k_50k = 'B19001_010',
          inc_btw_50k_60k = 'B19001_011',
          inc_btw_60k_75k = 'B19001_012',
          inc_btw_75k_100k = 'B19001_013',
          inc_btw_100k_125k = 'B19001_014',
          inc_btw_125k_150k = 'B19001_015',
          inc_btw_150k_200k = 'B19001_016',
          inc_gt_200k = 'B19001_017'
          )

census <- get_acs(geography = 'tract',
                 state = 'UT',
                 county = c('Salt Lake', 'Tooele'),
                 variables = vars,
                 output = 'wide',
                 geometry = TRUE)

```
#Combine income variables to create three new income categories.
We had retreived census data on the total number of households earning incomes 
in various income categories at the census block level. 
We then collapsed and determined the tract level data for that 
and combined all the income categories to form three broad categories 
These are under $35,000, between $35,000 and $60,000
and over $60,000.

```{r}
census <- census %>% 
  mutate(inc_quint_20 = inc_lt_10kE+inc_btw_10k_15kE+
          inc_btw_15k_20kE+inc_btw_20k_25kE+inc_btw_25k_30kE+inc_btw_30k_35kE)%>%
  mutate(inc_quint_35 = inc_btw_35k_40kE + inc_btw_40k_45kE 
        + inc_btw_45k_50kE + inc_btw_50k_60kE)%>%
  mutate(inc_quint_60 = inc_btw_60k_75kE + inc_btw_75k_100kE)%>%
  mutate(inc_quint_80 = inc_btw_100k_125kE+inc_btw_125k_150kE)%>%
  mutate(inc_quint_100 = inc_btw_150k_200kE+inc_gt_200kE)%>%
select(GEOID, 
         total_hhsE,
         no_vehE,
         avg_hh_sizeE,
         hh_1personE,
         hh_2personE,
         hh_3personE,
         hh_4person_plusE,
        
         inc_quint_20,
         inc_quint_35,
         inc_quint_60,
         inc_quint_80,
         inc_quint_100,
        
          inc_lt_10kE,
          inc_btw_10k_15kE,
          inc_btw_15k_20kE,
          inc_btw_20k_25kE,
          inc_btw_25k_30kE,
          inc_btw_30k_35kE,
          inc_btw_35k_40kE,
          inc_btw_40k_45kE,
          inc_btw_45k_50kE,
          inc_btw_50k_60kE,
          inc_btw_60k_75kE,
          inc_btw_75k_100kE,
          inc_btw_100k_125kE,
          inc_btw_125k_150kE,
          inc_btw_150k_200kE,
          inc_gt_200kE)
```
#Download block level employment data for Utah from the LEHD dataset.
For this exercise, we first retreived employment data from the
Longitudinal Employer-Household Dynamica (LEHD) dataset. 
We pulled census block-level data for the state of Utah from the 
Workplace Characteristics section for 2019. 
Variables we pulled include:
- Total employment or number of workers
- Basic employment which we added (includes Agriculture, Forestry,
Fishing, and Hunting (CNS01), Mining and extraction (CNS02), Utilities
(CNS03), Construction (CNS04), Manufacturing (CNS05), Wholesale
trade (CNS06) and Transportation and warehousing (CNS06).
- Retail employment
- Service employment (total - basic - retail)

```{r}

library(tidyverse)

lehd_blocks <- read_csv('https://lehd.ces.census.gov/data/lodes/LODES7/ut/wac/ut_wac_S000_JT00_2019.csv.gz') %>%
  rename(total_emp = C000) %>%
  mutate(basic_emp = CNS01+CNS02+CNS03+CNS04+CNS05+CNS06+CNS08+CNS09) %>%
  rename(retail_emp = CNS07) %>%
  mutate(service_emp = total_emp - basic_emp - retail_emp) %>%
  select(w_geocode, total_emp, basic_emp, retail_emp, service_emp)
```

#Collapse block level employment data to tract level data for Utah.
We then collapse the block level data to determined 
employment characteristics at the tract level.
```{r}
library(tidyverse)

lehd_tracts <- lehd_blocks %>%
  mutate(w_geocode = as.character(w_geocode)) %>%
  mutate(GEOID = substr(w_geocode, 1, 11)) %>%
  select(-w_geocode) %>%
  group_by(GEOID) %>%
  summarize(across(everything(), ~sum(.)))
```

#Join tract level employment data with population data 
We then combine employment data for Utah state with the 
previously compiled population data tracts within Salt Lake MSA boundary. 
This filters down the state level employment data to data 
for our two selected counties in the Salt Lake MSA area

```{r}
zones <- left_join(census, lehd_tracts)

```

#Separate zone geometry or boundaries from the zone dataset
```{r}
zone_boundaries <- zones %>%
  select(GEOID, geometry)

zone_data <- zones %>%
  st_drop_geometry()
```

#Summarize tract statistics
```{r}
#
table <- zone_data %>%
  pivot_longer(cols=c(total_hhsE,no_vehE,
         inc_quint_20,
         inc_quint_35,
         inc_quint_60,
         inc_quint_80,
         inc_quint_100,
          total_emp, basic_emp, retail_emp, service_emp,
         hh_1personE, hh_2personE, hh_3personE, hh_4person_plusE),
               names_to = "Variable",
               values_to = "value") %>%
  group_by(Variable) %>%
  summarize(Average = mean(value, na.rm=TRUE),
            `Standard deviation` = sd(value, na.rm=TRUE),
            Median = median(value, median(value, na.rm=TRUE)))

write_csv(zone_data, "images/zone_summary.csv", append = FALSE)

```

# Regional Data
```{r}
vars <- c(inc_lt_10k = 'B19001_002',
          inc_btw_10k_15k = 'B19001_003',
          inc_btw_15k_20k = 'B19001_004',
          inc_btw_20k_25k = 'B19001_005',
          inc_btw_25k_30k = 'B19001_006',
          inc_btw_30k_35k = 'B19001_007',
          inc_btw_35k_40k = 'B19001_008',
          inc_btw_40k_45k = 'B19001_008',
          inc_btw_45k_50k = 'B19001_010',
          inc_btw_50k_60k = 'B19001_011',
          inc_btw_60k_75k = 'B19001_012',
          inc_btw_75k_100k = 'B19001_013',
          inc_btw_100k_125k = 'B19001_014',
          inc_btw_125k_150k = 'B19001_015',
          inc_btw_150k_200k = 'B19001_016',
          inc_gt_200k = 'B19001_017')



region_inc <- get_acs(
  geography = "metropolitan statistical area/micropolitan statistical area",
  variables = vars,
  summary_var = 'B19001_001',
  geometry = FALSE) %>%
  filter(GEOID == "41620") %>%
  mutate(pct = estimate / summary_est) %>%
  mutate(cumul_pct = cumsum(pct)) %>%
  select(variable, cumul_pct)
```
```{r}
kable(region_inc, digits=2)
```

#Creating chloropleth maps for household, income and employment data
```{r}
total_hh_map <- ggplot(census) +
  geom_sf(aes(fill = total_hhsE), color="NA")

map_palette <- brewer.pal(5, "PuBuGn")

total_hh_map + 
theme_void(base_size = 8) +
scale_fill_gradientn(colors = map_palette,  name = "Total number of households") 

ggsave("images/total_hh_map.png",
       height = 4.25, width = 6.5, units = "in")
```

```{r}
no_veh_map <- ggplot(census) +
  geom_sf(aes(fill = no_vehE), color="NA")

map_palette <- brewer.pal(5, "PuBuGn")

no_veh_map + 
theme_void(base_size = 8) +
scale_fill_gradientn(colors = map_palette,  name = "Total number of zero-vehicle households") 

ggsave("images/no_veh_map.png",
       height = 4.25, width = 6.5, units = "in")
```

```{r}
total_emp_map <- ggplot(zones) +
  geom_sf(aes(fill = total_emp), color="NA")

map_palette <- brewer.pal(5, "PuBuGn")

total_emp_map + 
theme_void(base_size = 8) +
scale_fill_gradientn(colors = map_palette,  name = "Total number of employees") 

ggsave("images/total_emp_map .png",
       height = 4.25, width = 6.5, units = "in")
```

```{r}
basic_emp_map <- ggplot(zones) +
  geom_sf(aes(fill = basic_emp), color="NA")

map_palette <- brewer.pal(5, "PuBuGn")

basic_emp_map + 
theme_void(base_size = 8) +
scale_fill_gradientn(colors = map_palette,  name = "Total number of basic employees") 

ggsave("images/basic_emp_map.png",
       height = 4.25, width = 6.5, units = "in")
```


```{r}
retail_emp_map <- ggplot(zones) +
  geom_sf(aes(fill = retail_emp), color="NA")

map_palette <- brewer.pal(5, "PuBuGn")

retail_emp_map + 
theme_void(base_size = 8) +
scale_fill_gradientn(colors = map_palette,  name = "Total number of retail employees") 

ggsave("images/retail_emp_map.png",
       height = 4.25, width = 6.5, units = "in")
```
```{r}
service_emp_map <- ggplot(zones) +
  geom_sf(aes(fill = service_emp), color="NA")

map_palette <- brewer.pal(5, "PuBuGn")

service_emp_map + 
theme_void(base_size = 8) +
scale_fill_gradientn(colors = map_palette,  name = "Total number of service employees") 

ggsave("images/service_emp_map.png",
       height = 4.25, width = 6.5, units = "in")
```
# Generate treemap diagrams for income and household size composition of study area
```{r}
income_buckets <- c(sum(census$inc_quint_20)/sum(census$total_hhsE),
                    sum(census$inc_quint_35)/sum(census$total_hhsE),
                    sum(census$inc_quint_60)/sum(census$total_hhsE),
                    sum(census$inc_quint_80)/sum(census$total_hhsE),
                    sum(census$inc_quint_100)/sum(census$total_hhsE)                    
                    ) 
income_tibble <- tibble(bucket = c("Quantile 20%: Below 35k",
                               "Quantile 35%: 35-60k",
                               "Quantile 60%: 60-100k",
                               "Quantile 80%: 100-150k",
                               "Quantile 100%: Above 150k"),

                     Percentage = as.numeric(income_buckets*100)
                     )

ggplot(income_tibble, aes(area = Percentage, fill= bucket)) +
  geom_treemap(show.legend = FALSE) +
  geom_treemap_text(aes(label = paste(bucket, "\n",
                                      prettyNum(Percentage, digits = 0),
                                      "%",sep = "")), 
                        color = "white") +
  scale_fill_brewer(palette = "PuBuGn")

ggsave("images/treemap_income.png",
       height = 4.25, width = 6.5, units = "in")
```


```{r}
hhsize_buckets <- c(sum(census$hh_1personE)/sum(census$total_hhsE),sum(census$hh_2personE)/sum(census$total_hhsE),sum(census$hh_3personE)/sum(census$total_hhsE),sum(census$hh_4person_plusE)/sum(census$total_hhsE)) 
hhsize_tibble <- tibble(bucket = c("Households with 1 person",
                               "Households with 2 person",
                               "Households with 3 person",
                               "Households with 4 person and plus"),

                     Percentage = as.numeric(hhsize_buckets*100)
                     )

ggplot(hhsize_tibble, aes(area = Percentage, fill= bucket)) +
  geom_treemap(show.legend = FALSE) +
  geom_treemap_text(aes(label = paste(bucket, "\n",
                                      prettyNum(Percentage, digits = 0),
                                      "%",sep = "")), 
                        color = "white") +
  scale_fill_brewer(palette = "PuBuGn")

ggsave("images/treemap_hhsize.png",
       height = 4.25, width = 6.5, units = "in")
```

# Create Dot-Density Map
```{r}
library(sf)

hh_1person_pts <- st_sample(census, 
                            size = ceiling(census$hh_1personE/100))

hh_2person_pts <- st_sample(census, 
                            size = ceiling(census$hh_2personE/100))

hh_3person_pts <- st_sample(census, 
                            size = ceiling(census$hh_3personE/100))

hh_4person_pts <- st_sample(census, 
                            size = ceiling(census$hh_4person_plusE/100))
```


#Save and write datasets to a file.
```{r}
st_write(zone_boundaries, "zones/boundaries.geojson", append = FALSE)
write_csv(zone_data, "existing/data/zone_data.csv", append = FALSE)
write_csv(zone_data, "alternative/data/zone_data.csv", append = FALSE)
write_csv(region_inc, "existing/data/regional_income_quintile.csv", append = FALSE)
```
```{r}
library(leaflet)
library(htmlwidgets)

map <- leaflet(census) %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(popup = ~GEOID,
              weight = 1,
              opacity = 1,
              highlightOptions =
                highlightOptions(fillColor = "red"))

saveWidget(map, file = 'zones/boundaries.html')
```



